---
title: "Initial processing of glycoproteomics datasets"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: united
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

# General settings

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  include = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  eval = TRUE,
  fig.width = 4 * (1 + sqrt(5)) / 2, 
  fig.height = 4,
  dpi = 700
)
knitr::opts_knit$set(root.dir = "/Users/burtsche/Documents/01_repos/Glycoproteomics/")
```

## Packages

```{r, message=F, warning =F, include=}
library(tidyverse)
library(ggplot2); theme_set(cowplot::theme_cowplot(font_size = 15))
library("reshape2")
library(ggrepel)
library(knitr)
library(ggvenn)

mutate <- dplyr::mutate
select <- dplyr::select
group_by <- dplyr::group_by
```

```{r}
options(ggplot2.discrete.colour= c("orange3", "darkslateblue", "darkred", "darkgreen", "darkgrey", "darkviolet"))
options(ggplot2.discrete.fill= c("orange3", "darkslateblue", "darkred", "darkgreen", "darkgrey", "darkviolet"))
```

# Functions

```{r}

glyco_psm_processing <- function(path, dataset, species, version, qvalue_cutoff, purity_cutoff){
  
  psm <- read_tsv(path, col_types = cols()) %>% 
    mutate(dataset = dataset, species = species)
  
  # deal wit column name bugs
  colnames(psm) <- gsub(" ", "\\.", colnames(psm))
  
  if (version == "2") {
    print("Renaming columns of yet another MSFragger version")
    psm$Observed.Modifications_new <- psm$Observed.Modifications
    psm$Observed.Modifications <- psm$Total.Glycan.Composition
  }
  
  
  psm <- psm %>%
    as.data.frame() %>% 
    mutate(
      modification = ifelse(grepl("N\\(", Assigned.Modifications), "glycosylation", "unmodified"),
      glycosylation = str_extract_all(Assigned.Modifications, "\\d+N"),
      n_position = as.numeric(str_extract(glycosylation, "\\d+")) + Protein.Start - 1,
      # add mass bin information
      mass = str_extract(Observed.Modifications, " .+$"),
      mass = as.numeric(str_replace(mass, " % ", "")),
      # bin = cut_interval(mass, n = 10),
      bin = cut(mass, breaks = seq(0, 8000, 500))
      # add glycan group information
    ) %>% 
    separate(Observed.Modifications, into = c("glycan_content", "modification_mass"), sep = " % ", remove = F) %>%
    separate(glycan_content, into = c("hexnac", "hexnac_content", "hex", "hex_content", "end"), sep = "\\(|\\)", extra = "merge", remove = FALSE) %>%
    mutate(
      hexnac_content = as.numeric(hexnac_content),
      hex_content = as.numeric(hex_content),
      end = ifelse(is.na(end), "", end)
    ) %>%
    mutate(glycan_type = case_when(
      grepl("Phospho", glycan_content) ~ "phospho",
      grepl("NeuAc|NeuGc", glycan_content) ~ "sialylated",
      grepl("Fuc", glycan_content) ~ "fucosylated",
      hexnac_content == 2 & hex_content > 3 & end == "" ~ "high mannose",
      hexnac_content == 2 & hex_content %in% 1:3 & end == "" ~ "paucimannose",
      hexnac_content <= 2 & hex == "" & end == "" ~ "small",
      TRUE ~ "complex/hybrid"
    )) %>%
    mutate(glycosylation = as.character(glycosylation)) 
  
  
  psm_glyco <- psm %>% 
    filter(modification == "glycosylation") %>% 
    filter(`Glycan.q-value` < qvalue_cutoff) %>% 
    filter(Purity > purity_cutoff)
  
  summary <- data.frame(n_proteins = length(unique(psm_glyco$Protein.ID)),
                        n_peptides = nrow(distinct(psm_glyco, Protein.ID, Modified.Peptide, Observed.Modifications)),
                        n_sites = nrow(distinct(psm_glyco, Protein.ID, n_position)),
                        dataset = dataset,
                        species = species
  )
  
  
  
  res <- list("psm" = psm_glyco,"summary" = summary, "psm_total" = psm)
  
}



```


#1. N-glyco


## LF N-glyco human single-shot data 

```{r, eval = F}
res_HEK_Hela_optimisation <- glyco_psm_processing(
  path = "data/High_energy_PGC+singleshots_human_psm.tsv",
  dataset = "HEK_HeLa_optimisation_singleshots",
  species = "human\ncell lines",
  version = "20",
  qvalue_cutoff = 0.05, 
  # for lf datasets dont filter for purity
  purity_cutoff = -1
)

res_HEK_Hela_optimisation$psm <- res_HEK_Hela_optimisation$psm %>%
  mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+")) %>% 
  filter(!grepl("lf_glyco_F", sample)) %>% 
mutate(experiment = str_extract(sample, "singleshot|SDS|450-2000|700-2000"),
         species = str_extract(sample,"HeLa|HEK"),
         replicate  = str_extract(sample,"\\d$")) %>% 
  mutate(experiment = ifelse(grepl("SDS", experiment), "SDS_lysis", experiment),
         experiment = ifelse(grepl("-", experiment), "scan_ranges", experiment),
         replicate = ifelse(replicate == 0, "1", replicate)) 

res_HEK_Hela_optimisation$psm_total <- res_HEK_Hela_optimisation$psm_total %>%
  mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+")) %>% 
  filter(!grepl("lf_glyco_F", sample)) %>% 
  mutate(experiment = str_extract(sample, "singleshot|SDS|450-2000|700-2000"),
         species = str_extract(sample,"HeLa|HEK"),
         replicate  = str_extract(sample,"\\d$")) %>% 
  mutate(experiment = ifelse(grepl("SDS", experiment), "SDS_lysis", experiment),
         experiment = ifelse(grepl("-", experiment), "scan_ranges", experiment),
         replicate = ifelse(replicate == 0, "1", replicate)) 

save(res_HEK_Hela_optimisation, file = paste0("data/",Sys.Date(), "_res_HEK_Hela_optimisation_psm_processed.RData"))
```

## LF N-glyco mouse Single-shot data

```{r, eval =F}

res_singleshot_mousebrain <- glyco_psm_processing(
  path = "data/High_energy_singleshots_mouse_brain_7ppm_psm.tsv",
  dataset = "LF_brains_singleshot",
  species = "mouse\nbrain",
  version = "20",
  qvalue_cutoff = 0.05, 
  # for lf datasets dont filter for purity
  purity_cutoff = -1
)

res_singleshot_mousebrain$psm <- res_singleshot_mousebrain$psm %>%
  mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+"), experiment = "singleshot") %>% 
  group_by(sample) %>% 
  mutate(replicate = cur_group_id()) %>% 
  ungroup()

res_singleshot_mousebrain$psm_total <- res_singleshot_mousebrain$psm_total %>%
  mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+"), experiment = "singleshot") %>% 
  group_by(sample) %>% 
  mutate(replicate = cur_group_id()) %>% 
  ungroup()

save(res_singleshot_mousebrain, file =paste0("data/", Sys.Date(),"_Mouse_brain_lf_singleshot_psm_processed.RData"))
```


## LF N-glyco human cell lines fractionated


```{r}
res_HEK_HeLa_lf_PGC <- glyco_psm_processing(
  "data/High_energy_PGC+singleshots_human_psm.tsv", 
  "HEK_HeLa_lf_PGC_frac", "human\ncell line", 
  "20", 
  0.05, 
  -1
  )

res_HEK_HeLa_lf_PGC$psm <-res_HEK_HeLa_lf_PGC$psm %>% 
   mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+")) %>% 
  filter(grepl("lf_glyco_F\\d", sample))

res_HEK_HeLa_lf_PGC$psm_total <-res_HEK_HeLa_lf_PGC$psm_total %>% 
   mutate(sample = str_extract(Spectrum, "CP_[[:alnum:]_-]+|glyco_[[:alnum:]_-]+")) %>% 
  filter(grepl("lf_glyco_F\\d", sample))

res_HEK_HeLa_lf_PGC$summary <- res_HEK_HeLa_lf_PGC$psm %>% 
  summarise(
    n_peptides = n_distinct(Protein.ID, Modified.Peptide, Observed.Modifications),
     n_sites = n_distinct(Protein.ID, n_position),
     n_proteins = n_distinct(Protein.ID)
  )
  
save(res_HEK_HeLa_lf_PGC, file = paste0("data/",Sys.Date(), "_HEK_HeLa_lf_PGC_psm_processed.RData"))
```


## LF N-glyco Mouse brain fractionated - high energy


```{r, eval=F}
#data/paper/psm_files/labelfree/Mouse_brain_allfiles_final_psm.tsv
res_mousebrain_fractionated_Nglyco <- glyco_psm_processing(
  path = "data/High_energy_PGC_mouse_brain_7ppm_psm.tsv", 
  dataset = "Mouse_brain_PGC_high_energy_7ppm",
  species = "mouse\nbrain", 
  version = "20", 
  qvalue_cutoff = 0.05, 
  purity_cutoff = -1
  )

save(res_mousebrain_fractionated_Nglyco, file = paste0("data/",Sys.Date(),"_mousebrain_fractionated_Nglyco_psm_processed.RData"))
  
```

## LF N-glyco Mouse brain fractionated - stepped HCD

```{r, eval=F}
#psm_species <- glyco_psm_processing("data/brain/Mouse_brain_glyco_allfiles_filtered_final.tsv", "LF_brains", "mouse\nbrain")
# glyco_psm_processing("data/Brain_steppedHCD_psm.tsv", "LF_brains_steppedHCD", "mouse\nbrain")
res_mousebrain_steppedHCD <- glyco_psm_processing(
  path = "data/Mouse_brain_stepped_reprocessed_Fragpipev20_psm.tsv", 
  dataset = "Mouse_brain_stepped_reprocessed_Fragpipev20",
  species = "mouse\nbrain", 
  version = "20", 
  qvalue_cutoff = 0.02, 
  purity_cutoff = -1
)

save(res_mousebrain_steppedHCD, file = paste0("data/",Sys.Date(),"_res_mousebrain_steppedHCD_psm_processed.RData"))
  
```

### Combine

```{r}
# load("data/paper/processed/2024-01-08_res_mousebrain_steppedHCD_psm_processed.RData")
# load("data/paper/processed/2024-01-08_mousebrain_fractionated_Nglyco_psm_processed.RData")

load("data/2024-05-02_mousebrain_fractionated_Nglyco_psm_processed.RData")
load("data/2024-04-26_res_mousebrain_steppedHCD_psm_processed.RData")

psm_mousebrain_combined <- bind_rows(
  res_mousebrain_steppedHCD$psm,
  res_mousebrain_fractionated_Nglyco$psm
)

psm_mousebrain_combined$species %>%  unique


psm_mousebrain_combined_total <- bind_rows(
  res_mousebrain_steppedHCD$psm_total %>%  mutate(glycosylation = as.character(glycosylation)),
  res_mousebrain_fractionated_Nglyco$psm_total %>%  mutate(glycosylation = as.character(glycosylation))
)

psm_mousebrain_combined_total$species %>%  unique

psm_mousebrain_combined %>%  distinct(Protein.ID, Modified.Peptide, Observed.Modifications) %>%  nrow

load("data/2024-05-02_res_HEK_Hela_optimisation_psm_processed.RData")
load("data/2024-05-02_HEK_HeLa_lf_PGC_psm_processed.RData")


psm_human_combined <- bind_rows(
  res_HEK_HeLa_lf_PGC$psm,
  res_HEK_Hela_optimisation$psm
) %>%  
  mutate(species = "human\ncell line")

psm_human_combined$species %>%  unique

psm_human_combined_total <- bind_rows(
  res_HEK_HeLa_lf_PGC$psm_total,
  res_HEK_Hela_optimisation$psm_total
) %>%  
  mutate(species = "human\ncell line")

psm_human_combined_total$species %>%  unique

res_species <- list()

res_species$psm <- bind_rows(
  psm_human_combined, psm_mousebrain_combined
)

res_species$psm_total <- bind_rows(
  psm_human_combined_total, psm_mousebrain_combined_total
)

res_species$psm$species %>%  unique

save(res_species,
     file = paste0(Sys.Date(), "_combined_lf_data_human_and_mouse.RData"))

```

## Quant N-glyco microbiome mouse brain

```{r, eval = F}
res_microbiome_mousebrain_Nglyco <-  glyco_psm_processing(
  path = "~/Documents/01_repos/Glycoproteomics/data/Microbiome_glyco_fragpipev20_10ppm_psm.tsv",
  dataset = "TMT_microbiome_mousebrain_Nglyco", 
  species = "mouse\nbrain", 
  version = "1",
  qvalue_cutoff = 0.05, 
  purity_cutoff = 0.5
  )

save(res_microbiome_mousebrain_Nglyco, file =paste0("data/paper/processed/",Sys.Date(),"_microbiome_mousebrain_Nglyco_psm_processed.RData"))
```

#2. O-glyco

## LF frac mousebrain

```{r, eval =F}
psm <- read_tsv("data/Mouse_brain_high_energy_O_glyco_psm.tsv")%>% 
    mutate(dataset = "O-glycosylation", species = "mouse\nbrain")

  # deal wit column name bugs
  colnames(psm) <- gsub(" ", "\\.", colnames(psm))

  psm_glyco <- psm %>%
   # remove N-glycosylation
   filter(!grepl("N.{1}T|N.{1}S", Peptide)) %>%
   filter(!(grepl("N.$", Peptide) & Next.AA %in% c("S", "T"))) %>%
   filter(
     !is.na(Observed.Modifications) &
       !is.na(Glycan.Score) &
       Observed.Modifications != "no target matches" &
       !(grepl("Mod1", Observed.Modifications)) &
       !(grepl("PeakApex", Observed.Modifications))
   ) %>%
   mutate(
     glycosylation = str_extract_all(Assigned.Modifications, "\\d+N"),
     n_position = as.numeric(str_extract(glycosylation, "\\d+")) + Protein.Start - 1,
     # add mass bin information
     mass = str_extract(Observed.Modifications, " .+$"),
     mass = as.numeric(str_replace(mass, " % ", "")),
     # bin = cut_interval(mass, n = 10),
     # bin = cut(mass, breaks = seq(100, plyr::round_any(max(mass), 500), 100))
     # add glycan group information
   ) %>%
   separate(Observed.Modifications, into = c("glycan_content", "modification_mass"), sep = " % ", remove = F) %>%
   separate(glycan_content, into = c("hexnac", "hexnac_content", "hex", "hex_content", "end"), sep = "\\(|\\)", extra = "merge", remove = FALSE) %>%
   mutate(
     hexnac_content = as.numeric(hexnac_content),
     hex_content = as.numeric(hex_content),
     end = ifelse(is.na(end), "", end)
   ) %>%
   mutate(glycan_type = case_when(
     grepl("Phospho", glycan_content) ~ "O-phospho",
     grepl("NeuAc", glycan_content) ~ "O-sialylated",
     grepl("Fuc", glycan_content) ~ "O-fucosylated",
     grepl("Sulf", glycan_content) ~ "O-sulfated",
     hexnac_content == 1 & hex == "" ~ "O-GlcNac",
     hexnac_content %in% c("", 1, 2) & hex_content >= 1 & end == "" ~ "O-hexose",
     TRUE ~ "O-hybrid"
   )) %>%
   dplyr::select(-n_position)
  

  
  summary <- psm_glyco %>%
  mutate(pepID = paste0(Protein.ID, "_", Modified.Peptide, "_", Observed.Modifications)) %>%
  dplyr::select(Protein.ID, pepID)%>%
  summarise(
    n_glycoproteins = n_distinct(Protein.ID),
    n_glycopeptides = n_distinct(pepID)
  ) %>% 
    mutate(dataset = "O-glycosylation",
                        species = "mouse\nbrain")
  
  
  res_mousebrain_fractionated_Oglyco <- list("psm" = psm_glyco, "summary" = summary)

 
save(res_mousebrain_fractionated_Oglyco, file = paste0("data/",Sys.Date(),"_mousebrain_fractionated_Oglyco_psm_processed.RData"))
```

## LF frac steppedHCD mousebrain

```{r, eval =F}
psm <- read_tsv("data/Mouse_brain_stepped_O-glyco_psm.tsv")%>% 
    mutate(dataset = "O-glycosylation_steppedHCD", species = "mouse\nbrain")

  # deal wit column name bugs
  colnames(psm) <- gsub(" ", "\\.", colnames(psm))

  psm_glyco <- psm %>%
    # remove N-glycosylation
    filter(!grepl("N.{1}T|N.{1}S", Peptide)) %>%
    filter(!(grepl("N.$", Peptide) & Next.AA %in% c("S", "T"))) %>%
       filter(
     !is.na(Observed.Modifications) &
      !is.na(Glycan.Score) &
       Observed.Modifications != "no target matches" &
       !(grepl("Mod1", Observed.Modifications)) &
       !(grepl("PeakApex", Observed.Modifications))
   ) %>%
    mutate(
      glycosylation = str_extract_all(Assigned.Modifications, "\\d+N"),
      n_position = as.numeric(str_extract(glycosylation, "\\d+")) + Protein.Start - 1,
      # add mass bin information
      mass = str_extract(Observed.Modifications, " .+$"),
      mass = as.numeric(str_replace(mass, " % ", "")),
      # bin = cut_interval(mass, n = 10),
      #bin = cut(mass, breaks = seq(100, plyr::round_any(max(mass), 500), 100))
      # add glycan group information
    ) %>%
    separate(Observed.Modifications, into = c("glycan_content", "modification_mass"), sep = " % ", remove = F) %>%
    separate(glycan_content, into = c("hexnac", "hexnac_content", "hex", "hex_content", "end"), sep = "\\(|\\)", extra = "merge", remove = FALSE) %>%
    mutate(
      hexnac_content = as.numeric(hexnac_content),
      hex_content = as.numeric(hex_content),
      end = ifelse(is.na(end), "", end)
    ) %>%
    mutate(glycan_type = case_when(
      grepl("Phospho", glycan_content) ~ "O-phospho",
      grepl("NeuAc", glycan_content) ~ "O-sialylated",
      grepl("Fuc", glycan_content) ~ "O-fucosylated",
      grepl("Sulf", glycan_content) ~ "O-sulfated",
      hexnac_content == 1 & hex == "" ~ "O-GlcNac",
      hexnac_content %in% c("", 1, 2) & hex_content >= 1 & end == "" ~ "O-hexose",
      TRUE ~ "O-hybrid"
    )) %>% 
    select(-n_position)
  

  
  summary <- psm_glyco %>%
  mutate(pepID = paste0(Protein.ID, "_", Modified.Peptide, "_", Observed.Modifications)) %>%
  dplyr::select(Protein.ID, pepID)%>%
  summarise(
    n_glycoproteins = n_distinct(Protein.ID),
    n_glycopeptides = n_distinct(pepID)
  ) %>% 
    mutate(dataset = "O-glycosylation_steppedHCD",
                        species = "mouse\nbrain")
  
  
  res_mousebrain_fractionated_Oglyco_steppedHCD <- list("psm" = psm_glyco, "summary" = summary)

 
save(res_mousebrain_fractionated_Oglyco_steppedHCD, file = paste0("data/",Sys.Date(),"_mousebrain_fractionated_steppedHCD_Oglyco_psm_processed.RData"))
```

### Combine

```{r}
load("data/2024-05-06_mousebrain_fractionated_Oglyco_psm_processed.RData")
load("data/2024-05-06_mousebrain_fractionated_steppedHCD_Oglyco_psm_processed.RData")

res_mousebrain_fractionated_Oglyco_combined <- list()

res_mousebrain_fractionated_Oglyco_combined$psm <- ttt <-  bind_rows(
  res_mousebrain_fractionated_Oglyco$psm,
  res_mousebrain_fractionated_Oglyco_steppedHCD$psm
)

res_mousebrain_fractionated_Oglyco_combined$psm$species %>%  unique
res_mousebrain_fractionated_Oglyco_combined$psm$dataset%>%  unique

save(res_mousebrain_fractionated_Oglyco_combined, file = paste0("data/",Sys.Date(),"_mousebrain_fractionated_Oglyco_combined_psm_processed.RData"))
```

# Table fo Martin

```{r, eval = F}
load("2024-05-02_combined_lf_data_human_and_mouse.RData")
res_species$psm %>% 
  distinct(species, Protein.ID, Modified.Peptide, Peptide, n_position, glycan_content, glycan_type) %>% 
  write_csv("2024-05-03_glycotable_Martin.csv")
```



