---
title: "Preprocess tissue atlas data"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: united
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

# General settings

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  include = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  eval = TRUE,
  fig.width = 4 * (1 + sqrt(5)) / 2, 
  fig.height = 4,
  dpi = 700
)
knitr::opts_knit$set(root.dir = "/Users/burtsche/Documents/01_repos/Glycoproteomics/")
```

## Packages

```{r, message=F, warning =F}
library(tidyverse)
library(ggplot2);theme_set(cowplot::theme_cowplot(font_size = 15))
library("reshape2")
library(OmnipathR)
library(viridis)
library(ggrepel)
options(connectionObserver = NULL)
library(org.Mm.eg.db)
library(knitr)
library(ggpubr)
library(RColorBrewer)
library(ReactomePA)
library(clusterProfiler)
library(reactome.db)
library(ggvenn)
library(Biostrings)
library(limma)
library(ComplexHeatmap)
 library(nVennR)

mutate <- dplyr::mutate
select <- dplyr::select
group_by <- dplyr::group_by
filter <- dplyr::filter
```

```{r}
options(ggplot2.discrete.colour= c("darkorange", "darkblue", "darkgreen"),
        ggplot2.discrete.fill= c("darkorange", "darkblue", "darkgreen"))

order <- c("sialylated", "fucosylated", "high mannose", "complex/hybrid", "phospho", "paucimannose", "small")
pal <- c("lightgrey", "#E41A1C", "#377EB8", "#4DAF4A","#984EA3" ,"#FF7F00", "#FFFF33", "#A65628", "mediumseagreen", "grey")
names(pal) <- c("FP", "sialylated", "fucosylated", "high mannose", "complex/hybrid", "phospho", "paucimannose", "small", "hit", "no hit")

tilted <-  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust = 1))
```

# Functions

```{r}
source("~/Documents/generic_functions.R")
```

```{r, eval = F}
#load("results/2024-01-11_res_mouse_tissue_atlas.RData")
```


# 1. Glyco

## A Preprocess

```{r}
res_mouse_tissue_atlas <- glyco_psm_processing(
  path = "psm/Mouse_tissues_glyco_reprocessed_fragpipe20_10ppm_psm.tsv",
  dataset = "mouse_tissue_atlas_TMT",
  species = "Mus musculus",
  version = "20",
  qvalue_cutoff = 0.05, 
  purity_cutoff = 0.5
)

names(res_mouse_tissue_atlas) <- paste0("glyco_", names(res_mouse_tissue_atlas))
```

### TMT anno

```{r}
res_mouse_tissue_atlas$TMT_info <- data.frame(
  sample_glyco = c(
    "sample-01", "sample-02", "sample-03",
    "sample-04", "sample-05", "sample-06",
    "sample-07", "sample-08",
    "sample-09", "sample-10", "sample-11",
    "sample-12", "sample-13", "sample-14",
    "sample-15", "sample-16", "sample-17",
    "sample-18"
  ),
  replicate = c(
    rep("rep1", 3), rep("rep2", 3),  rep("rep3", 3),
    rep("rep1", 3), rep("rep2", 3),  rep("rep3", 3) 
  ),
  tissue = c(
    rep(c("brain", "kidney", "liver"), 6)
  ),
  mouse = c(
    rep(c("mouse1"), 9), 
    rep(c("mouse2"), 9)
  )
) %>% 
  mutate(group = paste(tissue, mouse, replicate, sep = "_"))
```

Define a long object, reused for many operations below

```{r}
res_mouse_tissue_atlas$glyco_psm_long <- res_mouse_tissue_atlas$glyco_psm %>%
  select(Protein.ID, Gene, Peptide, Modified.Peptide, Observed.Modifications, n_position, glycan_type, Gene, Purity, matches("sample"))%>%
  melt(
    variable.name = "sample",
    value.name = "quant",
    id.vars = c("Protein.ID", "Gene", "Peptide", "Modified.Peptide","Observed.Modifications", "n_position", "glycan_type", "Purity")
  ) %>%
  # add TMT info
  inner_join(res_mouse_tissue_atlas$TMT_info,
    by = c("sample" = "sample_glyco"),
    multiple = "all"
  ) 

```


### Noise filtering

```{r}
# calculate ratios from all channel to highest channel
res_mouse_tissue_atlas$noise_ratios <-tt <-  res_mouse_tissue_atlas$glyco_psm_long %>% 
  mutate(ID = paste0(Protein.ID, "_", Gene, "_", Modified.Peptide, "_", Observed.Modifications, "_", glycan_type, "_", n_position))%>%
  group_by(ID,group, tissue, replicate) %>%
  summarise(med_quant = median(quant)) %>%
  group_by(ID) %>%
  mutate(max_q = max(med_quant)) %>%
  mutate(log2fc = log2(med_quant / max_q))

res_mouse_tissue_atlas$noise_ratios %>%
  ggplot(aes(x = log2fc)) +
  geom_histogram()

# define what should be set to 0
res_mouse_tissue_atlas$noise <- tt <- res_mouse_tissue_atlas$noise_ratios %>%
  ungroup() %>% 
  filter(med_quant > 0) %>%
  filter(log2fc < -4) %>%
  select(ID, group) %>%
  mutate(noiseID = paste(ID, group, sep = "_"))
```

### check tissue ids

```{r}
# define in how many tissues a glycopeptide has been detected
res_mouse_tissue_atlas$tissue_ids <-tt<- res_mouse_tissue_atlas$glyco_psm_long %>% 
  drop_na() %>%
  mutate(
    ID = paste0(Protein.ID,"_", Gene, "_", Modified.Peptide, "_", Observed.Modifications, "_", glycan_type, "_", n_position),
    noiseID = paste(ID, group, sep = "_")
  ) %>%
  filter(!is.na(Protein.ID) & !is.na(Modified.Peptide) & !is.na(quant)) %>%
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
  mutate(quant = ifelse(quant == 0, NA, quant))%>%
  drop_na() %>% 
  dcast(ID ~ group, value.var = "quant", fun.aggregate = function(x) sum(x, na.rm =T))%>% 
  melt() %>% 
  filter(value > 0) %>%
  separate(variable, into = c("tissue","mouse", "rep")) %>% 
  group_by(ID, tissue, mouse) %>% 
  count%>% 
  # quant in all three replicates
  filter(n == 3) %>% 
  group_by(mouse, ID) %>% 
  count() %>% 
  separate(ID, into = c("Protein.ID","Gene", "Modified.Peptide", "Observed.Modifications","glycan_type", "n_position"), remove = F, sep = "_") %>% 
  mutate(n_position = as.character(n_position)) %>% 
  select(-ID) %>% 
  # rejoin with quant info
  inner_join(res_mouse_tissue_atlas$glyco_psm_long %>%  mutate(n_position = as.character(n_position))) %>% 
  rename(number_of_tissues = n) %>% 
  mutate(glycan_type = factor(glycan_type, levels = order))

```

```{r, eval = F}

input <- res_mouse_tissue_atlas$tissue_ids %>% 
  drop_na() %>% 
  ungroup() %>% 
  group_by(mouse) %>% 
  distinct(mouse, ID, Observed.Modifications, number_of_tissues) %>% 
  ungroup()
  
input2 <- input %>%  filter(mouse == "mouse1") %>%  select(-mouse, -Observed.Modifications)

v <- nVennR::plotVenn(split(input2 , input2$number_of_tissues),   opacity = 0.3, borderWidth = 3, nCycles = 50000)

input2 <- res_mouse_tissue_atlas$tissue_ids %>% 
  ungroup()%>%   
  distinct(tissue, Observed.Modifications) %>% 
  group_by(Observed.Modifications) %>% 
  count(name = "number_of_tissues") %>% 
  drop_na()

v <- nVennR::plotVenn(split(input2 , input2$number_of_tissues),   opacity = 0.3, borderWidth = 3, nCycles = 50000)
```


add mass info
add # sugars info

```{r, eval =F}
res_mouse_tissue_atlas$tissue_ids %>% 
  ungroup() %>% 
  distinct(mouse, glycan_type,number_of_tissues, ID) %>% 
  ggplot(aes(x = glycan_type)) +
  geom_bar(aes( fill = as.factor(number_of_tissues)), pos = "fill") +
  scale_fill_brewer(palette = "Reds") +
  geom_text(aes(label = after_stat(count)), stat = "count", position = "fill", vjust =-1, size =2)+
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust = 1)) +
  facet_wrap(~mouse)


res_mouse_tissue_atlas$tissue_ids %>% 
  ungroup() %>% 
  mutate(noiseID = paste(ID, group, sep = "_")) %>% 
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
  filter(number_of_tissues ==1) %>% 
  filter(quant >0 ) %>% 
  distinct(mouse, tissue, glycan_type,number_of_tissues, ID) %>% 
  ggplot(aes(x = tissue, fill = glycan_type)) +
  geom_bar(pos = "dodge") +
  scale_fill_manual(values = pal) +
  tilted +
  labs(title = "IDs of tissue specific glycopeptides", subtitle = "number_of_tissues = 1")+
  facet_wrap(~mouse)
```

#### number of complex sugars

```{r, eval =F}
res_mouse_tissue_atlas$tissue_ids %>%
 ungroup()  %>% 
  ungroup() %>% 
  mutate(noiseID = paste(ID, group, sep = "_")) %>% 
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
 filter(quant > 0) %>%
 distinct(mouse,number_of_tissues, ID, Observed.Modifications) %>%
 mutate(
   n_gc = str_extract(Observed.Modifications, "NeuGc\\(\\d+\\)"),
   n_gc = as.numeric(str_replace_all(n_gc, "NeuGc\\(|\\)", "")),
   n_ac = str_extract(Observed.Modifications, "NeuAc\\(\\d+\\)"),
   n_ac = as.numeric(str_replace_all(n_ac, "NeuAc\\(|\\)", ""))
 ) %>%
 rowwise() %>%
 mutate(
   n_sia = sum(c(n_ac, n_gc), na.rm = T)
 ) %>%
 select(-n_gc, -n_ac) %>%
 ggplot(aes(x = factor(n_sia))) +
 geom_bar(aes(fill = as.factor(number_of_tissues)), position = "fill") +
   geom_text(aes(label = after_stat(count)), stat = "count", position = "fill", vjust =-1, size =2)+
  scale_fill_brewer(palette = "Reds")+
  facet_wrap(~mouse)

res_mouse_tissue_atlas$tissue_ids %>%
 ungroup() %>% 
  ungroup() %>% 
  mutate(noiseID = paste(ID, group, sep = "_")) %>% 
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
 filter(quant > 0) %>%
 distinct(mouse, number_of_tissues, ID, Observed.Modifications) %>%
 mutate(
   n_fuc = str_extract(Observed.Modifications, "Fuc\\(\\d+\\)"),
   n_fuc = as.numeric(str_replace_all(n_fuc, "Fuc\\(|\\)", "")),
   n_fuc = ifelse(is.na(n_fuc), 0, n_fuc)
 ) %>%
 ggplot(aes(x = factor(n_fuc))) +
 geom_bar(aes(fill = as.factor(number_of_tissues)), position = "fill")+
  geom_text(aes(label = after_stat(count)), stat = "count", position = "fill", vjust =-1, size =2)+
  scale_fill_brewer(palette = "Reds")+
  facet_wrap(~mouse)
  
```


#### mass

```{r, eval =F}
res_mouse_tissue_atlas$tissue_ids %>% 
  ungroup()  %>% 
  ungroup() %>% 
  mutate(noiseID = paste(ID, group, sep = "_")) %>% 
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
  filter(quant >0) %>% 
  distinct(mouse, number_of_tissues, ID, Observed.Modifications) %>% 
  mutate(
  mass = str_extract(Observed.Modifications, " .+$"),
      mass = as.numeric(str_replace(mass, " % ", "")),
      # bin = cut_interval(mass, n = 10),
      bin = cut(mass, breaks = seq(0, 7000, 1000)),
      bin =as.factor(bin)) %>% 
  ggplot(aes(fill = bin, x = as.factor(number_of_tissues))) +
  geom_bar(pos = "dodge") +
  scale_fill_brewer(palette = "Reds") +
  facet_wrap(~mouse)
  
```

## B Normalisation

```{r}
enough_quant <-res_mouse_tissue_atlas$tissue_ids%>%  
  # in all tissues
  filter(number_of_tissues == 3)

res_mouse_tissue_atlas$glyco_psm_normalised <- 
  res_mouse_tissue_atlas$glyco_psm_long %>% 
  drop_na() %>%
  mutate(
    ID = paste0(Protein.ID,"_", Gene, "_", Modified.Peptide, "_", Observed.Modifications, "_", glycan_type, "_", n_position),
     noiseID = paste(ID, group, sep = "_"),
  ) %>%
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise$noiseID, 0, quant)) %>% 
  filter(ID %in% enough_quant$ID) %>%
  mutate(quant = ifelse(quant == 0, NA, quant)) %>%  
  drop_na() %>% 
  ungroup() %>% 
  group_modify(~ tibble(normalisation_melted_ptm(.))) %>%
  ungroup() 

rm(enough_quant)
```

### check

```{r}
ggplot(res_mouse_tissue_atlas$glyco_psm_normalised  %>%
         distinct(group, replicate,tissue,Protein.ID, Peptide, `Modified.Peptide`,Observed.Modifications, .keep_all = T), 
       aes(x = as.factor(sample), y = log2(quant), colour = tissue)) +
  geom_boxplot() +
  cowplot::theme_cowplot() +
  #facet_wrap(~replicate) +
  labs( y = "log2 intensity")

ggplot(res_mouse_tissue_atlas$glyco_psm_normalised  %>%
         distinct(group, replicate,tissue, Peptide, `Modified.Peptide`,Observed.Modifications, .keep_all = T), 
       aes(x = as.factor(sample), y = (quant_norm), colour = tissue)) +
  geom_boxplot() +
  cowplot::theme_cowplot() +
  #facet_wrap(~replicate) +
  labs( y = "log2 quant_norm")
```

## C PCA

### raw

```{r, eval =F}
m <- acast(res_mouse_tissue_atlas$glyco_psm_normalised %>% drop_na(),
  sample ~ Protein.ID + Modified.Peptide + Observed.Modifications,
  value.var = "quant",
  fun.aggregate = median,
  fill = 0
)

m[m == 0] <- NA
m <- m[complete.cases(m), ]

m <- t(m)
m <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(6000), ]
m <- t(m)

m <- log2(m)


pca <- prcomp(m)
plot(pca)
var = (pca$sdev)^2/sum(pca$sdev^2)*100

pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  inner_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>%
  ggplot(aes(x = PC1, y = PC2, colour =tissue, shape = paste0(replicate, "_", mouse))) +
  geom_point(size = 3) +
  scale_colour_brewer(palette = "Dark2") +
  labs(title = "Glycoproteome", 
       subtitle = "raw intensities", 
       x = paste0("PC1 (", as.character(round(var[1])), "%)"),
       y = paste0("PC2 (", as.character(round(var[2])), "%)"))

pca_rotation <- pca$rotation %>% as.data.frame()
```

### normalised

```{r, eval =F}
m <- acast(res_mouse_tissue_atlas$glyco_psm_normalised %>% drop_na(),
  sample ~ Protein.ID + Modified.Peptide + Observed.Modifications,
  value.var = "quant_norm",
  fun.aggregate = median,
  fill = 0
)

m[m == 0] <- NA
m <- t(m)
m <- m[complete.cases(m), ]


#m <- log2(m)

m <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(6000), ]

m <- t(m)

pca <- prcomp(m)
plot(pca)
var = (pca$sdev)^2/sum(pca$sdev^2)*100

pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  inner_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>%
  ggplot(aes(x = PC1, y = PC2, colour = as.factor(tissue), shape = paste0(mouse, "_", replicate))) +
  geom_point(size = 3) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Glycoproteome", 
       subtitle = "norm intensities", 
       x = paste0("PC1 (", as.character(round(var[1])), "%)"),
       y = paste0("PC2 (", as.character(round(var[2])), "%)"))

pca_rotation <- pca$rotation %>% as.data.frame()
```

```{r, eval =F}
rm(m, pca_rotation)
```

# vis example

```{r, eval =F}
res_mouse_tissue_atlas$glyco_psm_normalised %>% 
  filter(Protein.ID == "Q91ZV3" & n_position == 43) %>% 
  ggplot(aes(x = tissue, y = quant_norm, colour = tissue,group =ID)) +
  #geom_boxplot() +
  geom_jitter(position = position_dodge(width = 0.9)) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~mouse) +
  theme(axis.text.x = element_text(angle =45, hjust = 1, vjust =1))

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(Protein.ID == "Q91ZV3" & n_position == 43) %>% 
  ggplot(aes(x = tissue, y = cleaned_glycosignal, colour = tissue,group = glycoID)) +
  #geom_boxplot() +
  geom_jitter(position = position_dodge(width = 0.9)) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~mouse) +
  theme(axis.text.x = element_text(angle =45, hjust = 1, vjust =1))
```

#2. FP

## A Load and format

```{r}
res_mouse_tissue_atlas$FP <- read_tsv("psm/Tissues_FP_top2000_protein.tsv")
colnames(res_mouse_tissue_atlas$FP) <- gsub(" ", "\\.", colnames(res_mouse_tissue_atlas$FP))

res_mouse_tissue_atlas$FP <- res_mouse_tissue_atlas$FP%>% 
  filter(grepl("Mus", Organism) & !(grepl("contam", Protein)))

res_mouse_tissue_atlas$FP_psm <- read_tsv("psm/Tissues_FP_top2000_psm.tsv")
colnames(res_mouse_tissue_atlas$FP_psm) <- gsub(" ", "\\.", colnames(res_mouse_tissue_atlas$FP_psm))

res_mouse_tissue_atlas$FP_psm <- res_mouse_tissue_atlas$FP_psm%>%
  filter(Protein.ID %in% res_mouse_tissue_atlas$FP$Protein.ID)

```


### Noise filtering

```{r}
# calculate ratios from all channel to highest channel
res_mouse_tissue_atlas$noise_ratios_FP <-tt <-  res_mouse_tissue_atlas$FP %>% 
  filter(Unique.Peptides > 0) %>% 
  select(Protein.ID, Gene, matches(res_mouse_tissue_atlas$TMT_info$sample_glyco)) %>% 
  melt(variable.name = "sample", value.name = "quant") %>% 
  inner_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>% 
  filter(quant > 0 & is.finite(quant))%>% 
  drop_na()  %>% 
  mutate(ID = Protein.ID)%>%
  group_by(ID,group, tissue, replicate) %>%
  summarise(med_quant = median(quant)) %>%
  group_by(ID) %>%
  mutate(max_q = max(med_quant)) %>%
  mutate(log2fc = log2(med_quant / max_q))

res_mouse_tissue_atlas$noise_ratios_FP %>%
  ggplot(aes(x = log2fc)) +
  geom_histogram()

# define what should be set to 0
res_mouse_tissue_atlas$noise_FP <- tt <- res_mouse_tissue_atlas$noise_ratios_FP %>%
  ungroup() %>% 
  filter(med_quant > 0) %>%
  filter(log2fc < -4) %>%
  select(ID, group) %>%
  mutate(noiseID = paste(ID, group, sep = "_"))
```

## B Normalise

```{r}
res_mouse_tissue_atlas$FP_normalised <- tt <- res_mouse_tissue_atlas$FP %>% 
  filter(Unique.Peptides > 0) %>% 
  select(Protein.ID, Gene, matches(res_mouse_tissue_atlas$TMT_info$sample_glyco)) %>% 
  melt(variable.name = "sample", value.name = "quant") %>% 
  inner_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>% 
  filter(quant > 0 & is.finite(quant))%>% 
  drop_na() %>% 
  mutate(ID = Protein.ID)%>%
  mutate(noiseID = paste(ID, group, sep = "_"))%>% 
  mutate(quant = ifelse(noiseID %in% res_mouse_tissue_atlas$noise_FP$noiseID, 0, quant)) %>% 
  ungroup() %>% 
  group_modify(~ tibble(normalisation_melted_ptm(.))) %>%
  ungroup() 

res_mouse_tissue_atlas$FP_psm_normalised <- tt <- res_mouse_tissue_atlas$FP_psm %>% 
  select(Protein.ID, Gene, Peptide, Protein.Start, Protein.End, matches(res_mouse_tissue_atlas$TMT_info$sample_glyco)) %>% 
  melt(id.vars =c("Protein.ID", "Gene", "Peptide", "Protein.Start", "Protein.End"), variable.name = "sample", value.name = "quant") %>% 
  inner_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>% 
  filter(quant > 0 & is.finite(quant))%>% 
  drop_na() %>% 
  mutate(ID = paste0(Protein.ID,"_", Peptide))%>% 
  ungroup() %>% 
  group_modify(~ tibble(normalisation_melted_ptm(.))) %>%
  ungroup() 
```

### check

```{r, eval =F}
ggplot(res_mouse_tissue_atlas$FP_normalised, 
       aes(y = log2(quant), x = sample, colour = tissue)) +
  geom_boxplot() 

ggplot(res_mouse_tissue_atlas$FP_normalised, 
       aes(y = quant_norm, x = sample, colour = tissue)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust =1, vjust =1)) +
  labs(x = "")

```

## C PCA

### raw

```{r, eval =F}
m <- acast(res_mouse_tissue_atlas$FP_normalised %>% drop_na()%>%  filter(Protein.ID %in% res_mouse_tissue_atlas$glyco_psm$Protein.ID),
  sample ~ Protein.ID ,
  value.var = "quant",
  fun.aggregate = median,
  fill = 0
)

m[m == 0] <- NA
m <- t(m)
m <- m[complete.cases(m), ]
m <- t(m)

m <- log2(m)

m <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(1200), ]

pca <- prcomp(m)
plot(pca)
var = (pca$sdev)^2/sum(pca$sdev^2)*100

pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  left_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>%
  ggplot(aes(x = PC1, y = PC2, colour = as.factor(tissue), shape = replicate)) +
  geom_hline(yintercept = 0, linetype =3)+
  geom_vline(xintercept = 0, linetype =3)+
  geom_point(size = 3) +
  labs(title = "Proteome - glycoproteins", 
       subtitle = "raw intensities", 
       x = paste0("PC1 (", as.character(round(var[1])), "%)"),
       y = paste0("PC2 (", as.character(round(var[2])), "%)")) 

pca_rotation <- pca$rotation %>% as.data.frame()
```

### normalised

```{r, eval =F}
m <- acast(res_mouse_tissue_atlas$FP_normalised %>% drop_na(),
  sample ~ Protein.ID ,
  value.var = "quant_norm",
  fun.aggregate = median,
  fill = 0
)

m[m == 0] <- NA
m <- t(m)
m <- m[complete.cases(m), ]
m <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(1200), ]

m <- t(m)

pca <- prcomp(m)
plot(pca)
var = (pca$sdev)^2/sum(pca$sdev^2)*100

pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  left_join(res_mouse_tissue_atlas$TMT_info, by = c("sample" = "sample_glyco")) %>%
  ggplot(aes(x = PC1, y = PC2, colour = as.factor(tissue), shape = paste0(mouse, "_", replicate))) +
  geom_point(size = 3) +
  labs(title = "Proteome", 
       subtitle = "norm intensities", 
       x = paste0("PC1 (", as.character(round(var[1])), "%)"),
       y = paste0("PC2 (", as.character(round(var[2])), "%)"))

pca_rotation <- pca$rotation %>% as.data.frame()
```


```{r, eval =F}
rm(m, pca_rotation)
```



```{r}
res_mouse_tissue_atlas$FP_normalised %>% 
  filter(Protein.ID == "Q91ZV3") %>% 
  ggplot(aes(x = tissue, y = quant_norm, colour = tissue,group =ID)) +
  #geom_boxplot() +
  geom_jitter(position = position_dodge(width = 0.9)) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~mouse) +
  theme(axis.text.x = element_text(angle =45, hjust = 1, vjust =1))
```


# Enzyme abundance

```{r, eval =F}
enzymes <- data.frame(Gene =
                        c(
"FUT1",
"FUT3",
"FUT4",
"FUT6",
"FUT2",
"FUT5",
"FUT7",
"FUT11",
"FUT10",
"LFNG",
"FUT8",
"POFUT1",
"FUT9",
"ST8SIA5",
"ST8SIA3",
"ST6GAL1",
"ST8SIA6",
"ST3GAL1",
"ST3GAL3",
"ST3GAL4",
"ST3GAL2",
"ST6GALNAC3",
"ST8SIA1",
"ST8SIA2",
"ST8SIA4",
"ST6GALNAC6",
"ST6GAL2",
"ST6GALNAC5",
"ST6GALNAC4",
"ST6GALNAC1",
"ST6GALNAC2",
"ST3GAL5",
"ST3GAL6"
))

res_mouse_tissue_atlas$FP_normalised %>% 
  filter(toupper(Gene) %in% (enzymes$Gene)) %>% 
  ggplot(aes(x= reorder(Gene, quant_norm), y =quant_norm, colour = tissue)) +
  geom_point() +
  theme_bw() +
  tilted
```


# 3. Correct data

## Combine glyco and total

```{r}
length(unique(res_mouse_tissue_atlas$glyco_psm_normalised$Protein.ID))
length(unique(res_mouse_tissue_atlas$FP_normalised$Protein.ID))

length(intersect(unique(res_mouse_tissue_atlas$glyco_psm_normalised$Protein.ID), unique(res_mouse_tissue_atlas$FP_normalised$Protein.ID)))
```


```{r}
data <- res_mouse_tissue_atlas$glyco_psm_normalised %>%
  # format glyco data
  filter(quant >0) %>% 
  # remove psm duplicates which arise from normalisation
  distinct(ID,sample, quant_norm, .keep_all = T) %>% 
  dplyr::select(Protein.ID, glycoID = ID, sample, glyco_intensity = quant)%>%
  # add FP data
  inner_join(res_mouse_tissue_atlas$FP_normalised %>%
               filter(quant >0) %>% 
               #filter(quant_norm >0) %>% 
    dplyr::select(Protein.ID, sample, total_intensity = quant), by = c("sample", "Protein.ID"))
```

```{r}
data_corrected <- data %>%
  group_by(Protein.ID) %>%
  mutate(median_ptm = 2^median(log2(glyco_intensity + 1), na.rm = TRUE),
         median_proteomics = 2^median(log2(total_intensity + 1), na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(nominator_norm = (glyco_intensity / total_intensity) ) %>%
  mutate(denominator_factor = (median_ptm / median_proteomics) ) %>%
  mutate(norm_ptm_data = (nominator_norm / denominator_factor) * median_ptm) %>%
  dplyr::select(glycoID, norm_ptm_data, sample) %>%
  pivot_wider(names_from = sample, values_from = norm_ptm_data)%>% 
  column_to_rownames('glycoID') %>%
  as.matrix() %>%
  limma::normalizeVSN()
```

```{r}
data_corrected %>%
  cor(use = "pairwise.complete.obs", method = "spearman") %>%
  pheatmap()
```

```{r}
res_mouse_tissue_atlas$glyco_psm_corrected <- data_corrected %>% 
  as.data.frame() %>% 
  rownames_to_column("glycoID")%>% 
  melt(variable.name = "sample", value.name = "cleaned_glycosignal") %>%
  left_join(res_mouse_tissue_atlas$glyco_psm_normalised, by = c("glycoID" = "ID","sample"))
```

```{r}
rm(data, data_corrected)
```

```{r, eval =F}
res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  drop_na() %>% 
  acast(glycoID ~ tissue + replicate + mouse, value.var = "cleaned_glycosignal") %>% 
  cor(use = "pairwise.complete.obs", method = "spearman") %>% 
  ComplexHeatmap::Heatmap(col = circlize::colorRamp2(c(0,1), c( "white", "darkred")), name = "corr")

tt <- res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  drop_na() %>% 
  acast(glycoID ~ tissue + replicate + mouse, value.var = "cleaned_glycosignal") %>% 
  cor(use = "pairwise.complete.obs", method = "spearman")

tt %>%  
  as.data.frame() %>% 
  rownames_to_column("sample1") %>% 
  melt() %>% 
  separate(sample1, into = c("tissue1","replicate1", "mouse1")) %>% 
  separate(variable, into = c("tissue2","replicate2", "mouse2")) %>% 
  filter(tissue1 == tissue2) %>% 
  group_by(tissue1) %>% 
  summarise(m = mean(value))
```



# Investigate data

## PCA

```{r, eval =F}
# m <- acast(psm_glyco_brains_anno_corrected, sample ~ glycoID, value.var = "cleaned_glycosignal", fun.aggregate = median, fill = 0)
# m[is.na(m)] <- 0
m <- acast(res_mouse_tissue_atlas$glyco_psm_corrected, 
           glycoID ~ sample, value.var = "cleaned_glycosignal", fill = 0, fun.aggregate = median)
m[m== 0] <- NA
m <- m %>% 
  as.data.frame() %>% 
  drop_na()%>% 
  as.matrix()

m <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(6000), ]

m <- t(m)

pca <- prcomp(m)
plot(pca)
var = (pca$sdev)^2/sum(pca$sdev^2)*100

pca$x %>%
  as.data.frame() %>%
  rownames_to_column("sample_glyco") %>%
  left_join(res_mouse_tissue_atlas$TMT_info) %>%
  ggplot(aes(x = PC1, y = PC2, colour = tissue, shape = paste0(mouse, "_", replicate))) +
    geom_hline(yintercept = 0, linetype =3)+
  geom_vline(xintercept = 0, linetype =3)+
  geom_point(size = 3) +
  labs(title = "Glycoproteome", 
       subtitle = "normalized and corrected", 
       x = paste0("PC1 (", as.character(round(var[1])), "%)"),
       y = paste0("PC2 (", as.character(round(var[2])), "%)"))


rm(m, pca_rotation)
```

## heatmap

```{r, eval =F}
m <- acast(
  res_mouse_tissue_atlas$glyco_psm_corrected %>%  drop_na(),  
  glycoID~ tissue+mouse+replicate, value.var = "cleaned_glycosignal", 
  fill = 0, fun.aggregate = median)

m[m== 0] <- NA
m <- m %>% 
  as.data.frame() %>% 
  drop_na()%>% 
  as.matrix()

mat <- m[matrixStats::rowVars(m, na.rm = T) %>%
  order(decreasing = T) %>%
  head(6000), ]

```

```{r, eval =F}
library(cluster)
library(factoextra)

#create plot of number of clusters vs total within sum of squares
fviz_nbclust(scale(mat), kmeans, method = "wss")

# Use map_dbl to run many models with varying value of k
sil_width <- map_dbl(2:10,  function(k){
  model <- pam(x = scale(mat), k = k)
  model$silinfo$avg.width
})

# Generate a data frame containing both k and sil_width
sil_df <- data.frame(
  k = 2:10,
  sil_width = sil_width
)

# Plot the relationship between k and sil_width
ggplot(sil_df, aes(x = k, y = sil_width)) +
  geom_line() + geom_point() +
  scale_x_continuous(breaks = 2:10)
```

```{r, eval =F}
set.seed(1)
hm <- scale(mat) %>% ComplexHeatmap::Heatmap(show_row_names = F ,
                                      #row_km = 3, 
                                      use_raster =F, 
                                      name = "scaled glycointensity", 
                                      #col = circlize::colorRamp2(colors =brewer.pal("YlGnBu", n =4), breaks = c(10, 12, 14, 18)),
                                      show_row_dend = F )

hm = draw(hm); row_order(hm)
```


### clusters

```{r, eval =F}
rcl.list <- row_order(hm)  #Extract clusters (output is a list)

lapply(rcl.list, function(x) length(x))  #check/confirm size clusters

 # loop to extract genes for each cluster.
 for (i in 1:length(row_order(hm))){
 if (i == 1) {
 clu <- t(t(row.names(mat[row_order(hm)[[i]],])))
 out <- cbind(clu, paste("cluster", i, sep=""))
 colnames(out) <- c("ID", "Cluster")
 } else {
 clu <- t(t(row.names(mat[row_order(hm)[[i]],])))
 clu <- cbind(clu, paste("cluster", i, sep=""))
 out <- rbind(out, clu)
 }
 }

out <- out %>%  as.data.frame()%>% 
  separate(ID, into = c("Protein.ID", "Gene", "Modified.Peptide", "Observed.Modifications", "glycan_type", "glyco_position"), sep = "_", remove = F)
```

```{r, eval =F}
library(STRINGdb)
string_db <- STRINGdb$new( version="11.5", species=10090,
                           score_threshold=200, network_type="full", input_directory="")

example1_mapped <- res_mouse_tissue_atlas$glyco_psm %>%  ungroup()  %>%  distinct(Gene) %>% as.data.frame()
example1_mapped <- string_db$map(example1_mapped, "Gene", removeUnmappedRows = TRUE )

string_db <- STRINGdb$new( version="11.5", species=10090,
                           score_threshold=200, network_type="full", input_directory="",
                           backgroundV =  unique(example1_mapped$STRING_id))
```

```{r, eval = F}
tt <- out %>%
  distinct(Cluster, Gene) %>% 
  filter(toupper(Gene) %in% example1_mapped$Gene) %>% 
  drop_na() %>% 
  group_by(Cluster) %>%
  nest() %>%
  mutate(ER = map(data, function(df) {
    ER = string_db$get_enrichment(df$Gene)
    if(nrow(ER) >0){
       ER = as.data.frame(ER)
    } else {
      ER = data.frame(category = "dummy")
    }
  })) %>%
  select(-data) %>%
  unnest(cols = c(ER))

int <- tt %>%
  filter(category !=  "PMID") %>% 
  distinct(Cluster, description) %>% 
  group_by(description) %>% 
  mutate(n = n()) %>% 
  filter(n == 1)

tt %>% 
  filter(description %in% int$description) %>% 
  dplyr::arrange(Cluster)%>% 
  ggplot(aes(x = -log10(fdr), y = reorder(description, -log10(fdr)), colour = Cluster)) +
  geom_point(size =2)+
  facet_wrap(~Cluster, scales= "free") +
  theme(panel.grid.major.y = element_line(colour = "lightgray"), axis.text.y = element_text(size =6))
```

## glycan_types

```{r, eval = F}
res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  drop_na() %>% 
  ggplot(aes(x = tissue, colour = glycan_type, y = cleaned_glycosignal)) +
  geom_boxplot()

res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  ggplot(aes(x = tissue, colour = replicate, y = cleaned_glycosignal)) +
  geom_boxplot() +
  facet_wrap(~mouse)

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  drop_na() %>% 
  group_by(tissue,n_position, glycoID) %>% 
  summarise(mean_signal = median(2^cleaned_glycosignal)) %>% 
  group_by(glycoID, n_position) %>% 
  mutate(frac_signal = mean_signal/sum(mean_signal)) %>% 
  drop_na() %>% 
  separate(glycoID, into = c("Protein.ID", "Gene", "Modified.Peptide", "Observed.Modifications", "glycan_type", "glyco_position"), sep = "_", remove = F) %>% 
  ggplot(aes(x = tissue, colour = glycan_type, y = frac_signal)) +
  geom_boxplot() +
  tilted
```

## compositions

```{r}
top_ids <- res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(glycan_type %in% c("complex/hybrid", "fucosylated", "sialylated", "high mannose")) %>% 
  distinct(tissue, glycoID, Observed.Modifications ) %>% 
  group_by(tissue, Observed.Modifications) %>% 
  summarise(n = n()) %>% 
  group_by(tissue) %>% 
  slice_max(order_by = n, n = 10) %>% 
  arrange(n)

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(Observed.Modifications %in% top_ids$Observed.Modifications) %>% 
  ggplot(aes(x = (Observed.Modifications), y= cleaned_glycosignal, colour = tissue)) +
  geom_boxplot() +
  coord_flip()
```



# Correlations
## per site

```{r}
enough_ids <- res_mouse_tissue_atlas$glyco_psm_normalised %>% 
  distinct(Protein.ID,n_position, Modified.Peptide, Observed.Modifications) %>% 
  group_by(Protein.ID, n_position) %>% 
  count() %>% 
  filter(n >= 5) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position))

res_mouse_tissue_atlas$corr_per_site <- 
  res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(Protein.ID, Gene, n_position, Modified.Peptide, Observed.Modifications, group, mouse, tissue, replicate, cleaned_glycosignal, quant_norm) %>%
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  filter(siteID %in% enough_ids$siteID & cleaned_glycosignal >0) %>%
  group_by(Protein.ID, Gene, n_position) %>%
  nest() %>%
  mutate(mean_corr = map(data, function(df) {
    # prepare matrix composed of all psotions per protein
    t <- df %>%
      # melt(id.vars = "n_position") %>%
      acast(Modified.Peptide + Observed.Modifications ~ tissue +mouse, value.var = "cleaned_glycosignal", fun.aggregate = mean)
    
    t <- t[complete.cases(t),]
    # convert to correlation matrix
    t2 <- cor(t, method = "pearson")
    # remove correlation matrix as it messes with mean
    diag(t2) <- NA
    # back and forth conversion
    t3 <- as.data.frame(t2) %>%
      rownames_to_column("tissue1") %>%
      melt(id.vars = c("tissue1"), variable.name = "tissue2") %>%
      mutate(value = as.numeric(value))
  })) %>%
  select(-data) %>%
  unnest(cols = c(mean_corr))

res_mouse_tissue_atlas$corr_per_site_per_mouse <- 
  res_mouse_tissue_atlas$glyco_psm_corrected %>%
  distinct(Protein.ID, Gene, n_position, Modified.Peptide, Observed.Modifications, group, mouse, tissue, replicate, cleaned_glycosignal, quant_norm) %>%
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  filter(siteID %in% enough_ids$siteID & cleaned_glycosignal >0) %>%
  group_by(mouse, Protein.ID, Gene, n_position) %>%
  nest() %>%
  mutate(mean_corr = map(data, function(df) {
    # prepare matrix composed of all psotions per protein
    t <- df %>%
      # melt(id.vars = "n_position") %>%
      acast(Modified.Peptide + Observed.Modifications ~ tissue, value.var = "cleaned_glycosignal", fun.aggregate = mean)
    # convert to correlation matrix
    t2 <- cor(t, method = "pearson")
    # remove correlation matrix as it messes with mean
    diag(t2) <- NA
    # back and forth conversion
    t3 <- as.data.frame(t2) %>%
      rownames_to_column("tissue1") %>%
      melt(id.vars = c("tissue1"), variable.name = "tissue2") %>%
      mutate(value = as.numeric(value))
  })) %>%
  select(-data) %>%
  unnest(cols = c(mean_corr))

res_mouse_tissue_atlas$corr_per_site_per_mouse_gt <- tt <- 
  res_mouse_tissue_atlas$glyco_psm_corrected %>%
  distinct(Protein.ID, Gene, n_position, Modified.Peptide, Observed.Modifications, glycan_type, group, mouse, tissue, replicate, cleaned_glycosignal, quant_norm) %>%
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  filter(siteID %in% enough_ids$siteID & cleaned_glycosignal >0) %>%
  group_by(mouse,glycan_type, Protein.ID, Gene, n_position) %>%
  nest() %>%
  mutate(mean_corr = map(data, function(df) {
    # prepare matrix composed of all psotions per protein
    t <- df %>%
      # melt(id.vars = "n_position") %>%
      acast(Modified.Peptide + Observed.Modifications ~ tissue, value.var = "cleaned_glycosignal", fun.aggregate = mean)
    # convert to correlation matrix
    t2 <- cor(t, method = "pearson")
    # remove correlation matrix as it messes with mean
    diag(t2) <- NA
    # back and forth conversion
    t3 <- as.data.frame(t2) %>%
      rownames_to_column("tissue1") %>%
      melt(id.vars = c("tissue1"), variable.name = "tissue2") %>%
      mutate(value = as.numeric(value))
  })) %>%
  select(-data) %>%
  unnest(cols = c(mean_corr))
```

```{r}
res_mouse_tissue_atlas$corr_per_site_per_mouse %>% 
  drop_na() %>% 
  mutate(comparison = paste0(tissue1, "_", tissue2)) %>% 
  ggplot(aes(x = value, colour = mouse)) +
  geom_vline(xintercept = 0, linetype =3) +
  geom_density(size=0.8) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  facet_grid(tissue1~tissue2, scales ="free")

res_mouse_tissue_atlas$corr_per_site_per_mouse %>% 
  drop_na() %>% 
  distinct(mouse, value, .keep_all = T) %>% 
  mutate(comparison = paste0(tissue1, "_", tissue2)) %>% 
  ggplot(aes(x = value, colour = comparison)) +
  geom_vline(xintercept = 0, linetype =3) +
  geom_density(size=0.8) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  theme(panel.grid = element_blank()) 

res_mouse_tissue_atlas$corr_per_site %>% 
  drop_na() %>% 
  separate(tissue1, into = c("t1", "m1"), remove = F) %>% 
  separate(tissue2, into = c("t2", "m2"), remove = F) %>% 
  filter(t1 == t2) %>% 
  mutate(comparison = paste0(tissue1, "_", tissue2)) %>% 
  #filter(tissue1 == "liver_mouse1" & tissue2 == "kidney_mouse2") %>% 
  ggplot(aes(x = value, colour = t1)) +
  geom_vline(xintercept = 0, linetype =3) +
  geom_density(size=0.8) +
  scale_color_brewer(palette = "Dark2") +
  #facet_grid(t1~t2, scales ="free") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(title = "Corr of glycoforms per site in\nbtw same tissue of biological replicates")
```

```{r}
tt <- res_mouse_tissue_atlas$corr_per_site_per_mouse %>% 
  drop_na() %>% 
  mutate(comparison = paste0(tissue1, "_", tissue2)) %>% 
  dcast(Protein.ID + Gene + n_position + mouse ~ comparison, value.var = "value")
lowerFn <- function(data, mapping, method = "lm", ...) {
  p <- ggplot(data = data, mapping = mapping) +
    geom_abline(slope =1, colour = "grey") +
    geom_point(colour = "black", size = 0.5) +
    geom_smooth(method = method, color = "blue", ...)  +
    lims(x = c(-1, 1), y = c(-1, 1))
  p
}

tt %>% 
  GGally::ggpairs(columns = c(5,6,8),  lower = list(continuous = GGally::wrap(lowerFn, method = "lm") ))

```


## site-specific

```{r}
res_mouse_tissue_atlas$corr_per_site_per_mouse%>%
  filter(tissue1 != tissue2)%>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  group_by(mouse, Protein.ID, siteID) %>% 
  summarise(mean_corr = mean((value))) %>% 
  ungroup() %>% 
  summarise(m = mean(mean_corr))
```

```{r}
low_corr_sites <- res_mouse_tissue_atlas$corr_per_site_per_mouse %>% 
  filter(value <  0.6535837	) %>% 
   mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  ungroup() %>% 
  distinct(mouse, Protein.ID, siteID) %>% 
  mutate(type = "low_corr_site")

res_mouse_tissue_atlas$corr_per_site_per_mouse %>%  
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  mutate(low_corr_site_anno = ifelse(Protein.ID %in% low_corr_sites$Protein.ID, "with <0.7 site", "without <0.7 site")) %>% 
  ggplot(aes(x = value, colour = low_corr_site_anno)) +
  geom_density()

res_mouse_tissue_atlas$corr_per_site_per_mouse %>%  
  #filter(Protein.ID %in% low_corr_sites$Protein.ID) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  mutate(low_corr_site_anno = ifelse(Protein.ID %in% low_corr_sites$Protein.ID & siteID %in% low_corr_sites$siteID, "site <0.7", "site >0.7"))  %>% 
  mutate(low_corr_site_anno = ifelse(!(Protein.ID %in% low_corr_sites$Protein.ID), "other proteins", low_corr_site_anno))  %>% 
  ggplot(aes(x = value, colour = low_corr_site_anno)) +
  geom_density()

res_mouse_tissue_atlas$corr_per_site_per_mouse %>%  
  #filter(Protein.ID %in% low_corr_sites$Protein.ID) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>%
  mutate(low_corr_site_anno = ifelse(Protein.ID %in% low_corr_sites$Protein.ID & siteID %in% low_corr_sites$siteID, "site <0.7", "other"))  %>% 
  ggplot(aes(x = value, colour = low_corr_site_anno)) +
  geom_density()


res_mouse_tissue_atlas$corr_per_site_per_mouse %>%  
  group_by(mouse, Protein.ID, n_position) %>% 
  summarise(mean_corr_site = mean(value, na.rm =T)) %>% 
  group_by(mouse, Protein.ID) %>%
  mutate(n = n()) %>%
  filter(n > 2) %>%
  select(-n) %>% 
  group_by(mouse, Protein.ID) %>% 
  mutate(sum = sum(mean_corr_site), n_sites = n()) %>% 
  ungroup() %>% 
  drop_na() %>% 
  mutate(mean_corr_other_sites_on_protein = (sum-mean_corr_site)/(n_sites-1)) %>% 
  ggplot(aes(x = mean_corr_site, y = mean_corr_other_sites_on_protein)) +
  geom_abline(slope =1) +
  geom_point(size = 1, alpha = 0.5, stroke = NA) +
  #facet_wrap(~mouse) +
  lims(x=c(-0.6,1), y=c(-0.6,1)) +
  coord_fixed() +
  stat_smooth(aes(colour = mouse), method = "lm")

 res_mouse_tissue_atlas$corr_per_site_per_mouse %>%  
 group_by(mouse, Protein.ID) %>%
  mutate(n_sites_protein = n_distinct(n_position)) %>%
  filter(n_sites_protein > 2) %>% 
  #filter(tissue1 != tissue2)%>% 
  group_by(mouse, Protein.ID, n_position, n_sites_protein) %>% 
  summarise(mean_corr = mean(value, na.rm =T)) %>% 
  mutate(type = ifelse(mean_corr < 0.6691 , "low_corr_site", "other"))%>% 
  group_by(mouse,Protein.ID, n_sites_protein, type) %>% 
  summarise(n_sites_type = n()) %>% 
  ungroup() %>% 
  mutate(frac= n_sites_type/ n_sites_protein) %>% 
  filter(type == "low_corr_site") %>% 
  ggplot(aes(x=frac, fill = mouse, colour = mouse)) +
   geom_vline(xintercept = 0.5, linetype = 3) +
  geom_density(alpha = 0.2, size =1)+
  annotate("text", x=0.7, y=1.9, label="50% of sites\nshow low correlation\nacrsoss tissues") +
  #facet_wrap(~mouse) +
  labs(x = "fraction of low-correlation sites per protein")
```



# Domains

Retrieve domain annotations from interpro

```{r}
library(biomaRt)
ensembl <- useEnsembl(biomart = "genes", dataset = "mmusculus_gene_ensembl")

ipro = getBM(attributes=c("uniprot_gn_id", "interpro","interpro_description","interpro_short_description", "interpro_start", "interpro_end"), 
             filters = "uniprot_gn_id",
             values= unique(res_mouse_tissue_atlas$glyco_psm$Protein.ID), 
             mart=ensembl)

```

```{r, fig.height=6, fig.width=7}
#0.6691
# 0.666813
# 0.6545566	
hit_def <- res_mouse_tissue_atlas$corr_per_site_per_mouse%>%
  filter(tissue1 != tissue2)%>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  group_by(mouse, Protein.ID, siteID) %>% 
  summarise(mean_corr = mean((value))) %>% 
  mutate(hit =  ifelse(mean_corr > 0.6535837, "corr_up", "no trend"))%>% 
  mutate(mouse = "mouse") 

res_mouse_tissue_atlas$domain_ORA <- tt <- hit_def %>% 
  mutate(mouse = "mouse") %>% 
  ungroup() %>% 
  separate(siteID, into = c("Gene", "n_position"), sep = "_") %>% 
  inner_join(ipro,by = c("Protein.ID"="uniprot_gn_id"), multiple = "all") %>% 
  rename(corr_hit = hit) %>% 
  mutate(n_position = as.numeric(n_position)) %>% 
  mutate(in_domain = ifelse(n_position> interpro_start & n_position < interpro_end, TRUE, FALSE)) %>% 
  filter(!is.na(interpro_description) & !is.na(interpro_start)) %>% 
  select(mouse, Protein.ID, corr_hit, interpro_description, in_domain, n_position)%>% 
  filter(in_domain == T) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>% 
  mutate(n_sites = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit) %>% 
  mutate(n_glycan = length(unique(siteID))) %>% 
  group_by(mouse, interpro_description) %>% 
  mutate(n_domain = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit, interpro_description) %>% 
  mutate(n_type_domain = length(unique(siteID))) %>% 
  ungroup() %>% 
  distinct(mouse, corr_hit, interpro_description, n_sites,n_glycan, n_domain, n_type_domain) %>% 
  rowwise() %>% 
  mutate(FT = fisher.test(x = matrix(c(n_type_domain, n_glycan-n_type_domain, n_domain-n_type_domain, n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$p.value,
         odds_ratio = fisher.test(x = matrix(
           c(n_type_domain, 
             n_glycan-n_type_domain, 
             n_domain-n_type_domain,
             n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$estimate)%>% 
  mutate(adj.p.val = p.adjust(FT, method = "BH"),
         hit = ifelse(adj.p.val < 0.05, "hit (adj.p < 0.05)", "no hit (adj.p > 0.05)"))

hits <- res_mouse_tissue_atlas$domain_ORA  %>% 
  filter(hit == "hit (adj.p < 0.05)")

hits %>% 
  mutate(odds_ratio = ifelse(!is.finite(odds_ratio),20, odds_ratio))%>% 
  filter(n_type_domain > 10) %>% 
  ggplot(aes(y = interpro_description, x = log2(odds_ratio), shape = mouse,size = n_type_domain, colour = corr_hit)) +
  geom_point() +
  geom_vline(xintercept = c(-1.7, 0, 1.7), linetype = c(3,1,3)) +
  #scale_colour_manual(values = mycolors) +
  scale_size(range = c(3, 6)) +
  labs(y = "", subtitle = "adj.p.val < 0.05, more than 20 obs,mean corr cutoff,\nsame tissue comps removed") +
  theme_bw()

```

```{r}
res_mouse_tissue_atlas$corr_per_site_per_mouse %>% 
  filter(tissue1 != tissue2)%>% 
  inner_join(ipro,by = c("Protein.ID"="uniprot_gn_id"), multiple = "all") %>% 
  mutate(n_position = as.numeric(n_position)) %>% 
  mutate(in_domain = ifelse(n_position> interpro_start & n_position < interpro_end, TRUE, FALSE)) %>% 
  filter(!is.na(interpro_description) & !is.na(interpro_start)) %>% 
  mutate(domain= ifelse(grepl("Cadherin-like superfamily|Immunoglobulin V-set domain|Fibronectin type III superfamily|Immunoglobulin domain superfamily", interpro_description), interpro_description, "other")) %>% 
  ggplot(aes(y = value, x = domain, colour = domain)) +
  #ggforce::geom_sina() +
  geom_violin() +
  stat_summary(aes(group = domain),fun = median, geom = "crossbar",
               position = position_dodge(0.9), colour = "black") +
  theme(axis.text.x = element_blank()) +
  stat_compare_means(comparisons = list(c("Cadherin-like superfamily" , "other"), 
                                        c("Immunoglobulin V-set domain" , "other"),
                                        c("Fibronectin type III superfamily" , "other"))) 
  facet_wrap(~mouse)
  
```


```{r}
hit_def <- res_mouse_tissue_atlas$corr_per_site_per_mouse_gt%>%
  filter(tissue1 != tissue2)%>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  group_by(mouse, glycan_type, Protein.ID, siteID) %>% 
  summarise(mean_corr = mean((value))) %>% 
  mutate(hit =  ifelse(mean_corr > 0.6535837, "corr_up", "no trend"))

res_mouse_tissue_atlas$domain_ORA_gt <- tt <- hit_def %>% 
  ungroup() %>% 
  separate(siteID, into = c("Gene", "n_position"), sep = "_") %>% 
  inner_join(ipro,by = c("Protein.ID"="uniprot_gn_id"), multiple = "all") %>% 
  rename(corr_hit = hit) %>% 
  mutate(n_position = as.numeric(n_position)) %>% 
  mutate(in_domain = ifelse(n_position> interpro_start & n_position < interpro_end, TRUE, FALSE)) %>% 
  filter(!is.na(interpro_description) & !is.na(interpro_start)) %>% 
  select(glycan_type, mouse, Protein.ID, corr_hit, interpro_description, in_domain, n_position)%>% 
  filter(in_domain == T) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>% 
  mutate(n_sites = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit, glycan_type) %>% 
  mutate(n_glycan = length(unique(siteID))) %>% 
  group_by(mouse, interpro_description) %>% 
  mutate(n_domain = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit, interpro_description, glycan_type) %>% 
  mutate(n_type_domain = length(unique(siteID))) %>% 
  ungroup() %>% 
  distinct(glycan_type, mouse, corr_hit, interpro_description, n_sites,n_glycan, n_domain, n_type_domain) %>% 
  rowwise() %>% 
  mutate(FT = fisher.test(x = matrix(c(n_type_domain, n_glycan-n_type_domain, n_domain-n_type_domain, n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$p.value,
         odds_ratio = fisher.test(x = matrix(
           c(n_type_domain, 
             n_glycan-n_type_domain, 
             n_domain-n_type_domain,
             n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$estimate)%>% 
  mutate(adj.p.val = p.adjust(FT, method = "BH"),
         hit = ifelse(adj.p.val < 0.05, "hit (adj.p < 0.05)", "no hit (adj.p > 0.05)"))

hits <- res_mouse_tissue_atlas$domain_ORA_gt  %>% 
  filter(hit == "hit (adj.p < 0.05)")

hits %>% 
  mutate(odds_ratio = ifelse(!is.finite(odds_ratio),20, odds_ratio))%>% 
  filter(n_type_domain > 20) %>% 
  ggplot(aes(y = interpro_description, x = log2(odds_ratio), shape = mouse,size = n_type_domain, colour = corr_hit)) +
  geom_point() +
  geom_vline(xintercept =0) +
  #scale_colour_manual(values = mycolors) +
  #scale_size(range = c(2.5, 6)) +
  labs(y = "", subtitle = "adj.p.val < 0.05, at least 20 obs, 0.7 corr cutoff,\nsame tissue comps removed") +
  theme_bw() +
  facet_wrap(~glycan_type, nrow = 1)
```

```{r}
res_mouse_tissue_atlas$corr_per_site_per_mouse_gt %>% 
  filter(tissue1 != tissue2)%>% 
  inner_join(ipro,by = c("Protein.ID"="uniprot_gn_id"), multiple = "all") %>% 
  mutate(n_position = as.numeric(n_position)) %>% 
  mutate(in_domain = ifelse(n_position> interpro_start & n_position < interpro_end, TRUE, FALSE)) %>% 
  filter(!is.na(interpro_description) & !is.na(interpro_start)) %>% 
  mutate(domain= ifelse(grepl("Cadherin-like superfamily|Immunoglobulin-like fold|Fibronectin type III superfamily|Immunoglobulin domain superfamily", interpro_description), interpro_description, "other")) %>% 
  ggplot(aes(y = value, x = domain, colour = mouse)) +
  #ggforce::geom_sina() +
  geom_violin() +
  ggforce::geom_sina() +
  stat_summary(aes(group = domain),fun = median, geom = "crossbar",
               position = position_dodge(0.9), colour = "black") +
  #theme(axis.text.x = element_blank()) +
  # stat_compare_means(comparisons = list(c("Cadherin-like superfamily" , "other"), 
  #                                       c("Immunoglobulin-like fold" , "other"),
  #                                       c("Fibronectin type III superfamily" , "other"))) 
  facet_wrap(~glycan_type) +
  tilted

res_mouse_tissue_atlas$corr_per_site_per_mouse_gt %>% 
  filter(tissue1 != tissue2)%>% 
  drop_na() %>% 
  filter(glycan_type %in% c("fucosylated", "sialylated", "high mannose", "complex/hybrid")) %>% 
  #filter(glycan_type %in% c("phospho", "paucimannose", "small")) %>% 
  ggplot(aes(y = value, x = glycan_type, colour = glycan_type)) +
  ggforce::geom_sina(alpha = 0.2, position = position_dodge(width = 0.9)) +
  geom_boxplot(position = position_dodge(width = 0.9), fill = NA) +
    stat_summary(aes(group = glycan_type),fun = median, geom = "crossbar",
               position = position_dodge(0.9), colour = "black") +
  scale_colour_manual(values = pal) +
  tilted
```

```{r}
ids_per_site <- res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>% 
  drop_na() %>% 
  distinct(glycoID, siteID) %>% 
  group_by(siteID) %>% 
  count(name = "n_per_site")

hit_def <- res_mouse_tissue_atlas$corr_per_site_per_mouse %>%
  filter(tissue1 != tissue2)%>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  group_by(mouse, Protein.ID, siteID) %>% 
  summarise(mean_corr = mean((value))) %>% 
  mutate(hit =  ifelse(mean_corr > 0.6535837, "corr_up", "no trend")) %>% 
  ungroup() %>% 
  inner_join(ids_per_site)

hit_def %>% 
  filter(n_per_site < 200) %>% 
  ggplot(aes(y = (n_per_site), x = hit)) +
  geom_violin()


```



# Fractional intensities

```{r}
n_mod <-  res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(glycoID, Observed.Modifications) %>% 
  group_by(Observed.Modifications) %>% 
  count()

fractional_intensities <- res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(mouse, Observed.Modifications, tissue, replicate, cleaned_glycosignal) %>% 
  inner_join(n_mod) %>% 
  filter(n > 10) %>% 
  select(-n) %>% 
  mutate(cleaned_glycosignal = 2^cleaned_glycosignal) %>% 
  group_by(mouse, Observed.Modifications, tissue) %>% 
  summarise(cleaned_glycosignal = sum(cleaned_glycosignal)) %>% 
  group_by(mouse, Observed.Modifications) %>% 
  mutate(total_int = sum(cleaned_glycosignal)) %>% 
  ungroup() %>% 
  mutate(frac_int = cleaned_glycosignal/total_int)

top_comps <- fractional_intensities %>% 
  group_by(mouse, tissue) %>% 
  slice_max(order_by = frac_int, n = 10)

top_tissue <- fractional_intensities %>% 
  filter(Observed.Modifications %in% top_comps$Observed.Modifications) %>% 
  group_by(mouse, Observed.Modifications) %>% 
  mutate(top_tissue = ifelse(frac_int == max(frac_int), tissue, "not top")) %>% 
  ungroup() %>% 
  filter(top_tissue != "not top") %>% 
  distinct(mouse, Observed.Modifications, top_tissue)

# fractional_intensities %>% 
#   filter(Observed.Modifications %in% top_comps$Observed.Modifications) %>% 
#   inner_join(top_tissue) %>% 
#   filter(mouse == "mouse2") %>% 
#   mutate(Observed.Modifications= tidytext::reorder_within(Observed.Modifications, frac_int, top_tissue))%>%
#   ggplot(aes(x = Observed.Modifications, y = frac_int, fill = tissue)) +
#   geom_col(pos = "stack") +
#   facet_wrap(~top_tissue, scales="free")
#   geom_bar(position = "stack", stat = "identity") +
#   #tidytext::scale_x_reordered() +
#   # tilted +
#   # theme(axis.text.x = element_text(size =5)) +
#   facet_wrap(~top_tissue, scales="free")
```


```{r}
fractional_intensities <- res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(glycoID, mouse, Observed.Modifications, tissue, replicate, cleaned_glycosignal) %>% 
  mutate(cleaned_glycosignal = 2^cleaned_glycosignal) %>% 
  group_by(mouse, glycoID,Observed.Modifications, tissue) %>% 
  summarise(cleaned_glycosignal = sum(cleaned_glycosignal)) %>% 
  group_by(mouse, glycoID, Observed.Modifications) %>% 
  mutate(total_int = sum(cleaned_glycosignal)) %>% 
  ungroup() %>% 
  mutate(frac_int = cleaned_glycosignal/total_int)

top10 <- fractional_intensities %>% 
  distinct(mouse, glycoID, Observed.Modifications, tissue, frac_int)%>% 
  group_by(mouse, tissue) %>% 
  slice_max(prop = 0.01, order_by = frac_int) %>% 
  distinct(mouse, tissue, glycoID,Observed.Modifications, frac_int) %>% 
  ungroup() %>% 
  mutate(  mass = str_extract(Observed.Modifications, " .+$"),
      mass = as.numeric(str_replace(mass, " % ", ""))) %>% 
  mutate(
    fuc_cat = str_extract_all(Observed.Modifications, "Fuc\\(\\d+\\)"),
    sia_cat = str_extract_all(Observed.Modifications, "NeuGc\\(\\d+\\)|NeuAc\\(\\d+\\)")
  ) %>%
  mutate(
    fuc_cat = sapply(
      str_extract_all(fuc_cat, "[[:digit:]]+"),
      function(x) ifelse(identical(x, character(0)), NA, sum(as.numeric(x)))
    ),
    sia_cat = sapply(
      str_extract_all(sia_cat, "[[:digit:]]+"),
      function(x) ifelse(identical(x, character(0)), NA, sum(as.numeric(x)))
    )
  ) %>% 
  left_join(res_mouse_tissue_atlas$glyco_psm_corrected %>%  distinct(Observed.Modifications, glycan_type))

top10 %>% 
  ggplot(aes(x = mass, colour =tissue)) +
  geom_density() +
  #facet_wrap(~mouse) +
  labs(subtitle = "top1 frac int per tissue")


top10 %>% 
  mutate(tissue = factor(tissue, levels = c("brain","liver", "kidney"))) %>% 
  ggplot(aes(y = mass, x = tissue, colour =tissue)) +
  ggforce::geom_sina(stroke = NA, alpha = 0.2) +
  geom_boxplot(colour = "black", fill =NA) +
  stat_compare_means(comparisons = list(c("brain", "liver"), c("brain", "kidney"))) +
  #facet_wrap(~mouse) +
  labs(subtitle = "top1 frac int per tissue")

top10 %>% 
  ggplot(aes(x = tissue, fill =glycan_type)) +
  geom_bar()  +
  scale_fill_manual(values = pal)+
  #facet_wrap(~mouse)+
  labs(subtitle = "top1 frac int per tissue")

top10 %>% 
  filter(fuc_cat > 0) %>% 
  ggplot(aes(x = tissue, fill = as.factor(fuc_cat))) +
  geom_bar() +
  scale_fill_brewer(palette = "Reds")+
  labs(subtitle = "top1 frac int per tissue")

top10 %>% 
  filter(sia_cat > 0) %>% 
  ggplot(aes(x = tissue, fill = as.factor(sia_cat))) +
  geom_bar() +
  scale_fill_brewer(palette = "Reds")+
  labs(subtitle = "top1 frac int per tissue")

top10 %>% 
  mutate(sia_type = ifelse(grepl("NeuGc", Observed.Modifications), "NeuGc", "not sialylated"),
         sia_type = ifelse(grepl("NeuAc", Observed.Modifications), "NeuAc", sia_type),
         sia_type = ifelse(grepl("NeuAc", Observed.Modifications) & grepl("NeuGc", Observed.Modifications), "both", sia_type)) %>% 
  filter(sia_type != "not sialylated") %>% 
  ggplot(aes(x = tissue, fill =sia_type)) +
  geom_bar(pos = "fill")  +
  scale_fill_brewer(palette = "Dark2")+
  #facet_wrap(~mouse)+
  tilted +
  labs(subtitle = "top1 frac int per tissue", y = "relative count")

```

```{r}
fractional_intensities <- res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(mouse, glycan_type, Observed.Modifications, tissue, replicate, cleaned_glycosignal) %>% 
  inner_join(n_mod) %>% 
  filter(n > 10) %>% 
  select(-n) %>% 
  mutate(cleaned_glycosignal = 2^cleaned_glycosignal) %>% 
  group_by(mouse,glycan_type, Observed.Modifications, tissue) %>% 
  summarise(cleaned_glycosignal = sum(cleaned_glycosignal)) %>% 
  group_by(mouse,glycan_type, Observed.Modifications) %>% 
  mutate(total_int = sum(cleaned_glycosignal)) %>% 
  ungroup() %>% 
  mutate(frac_int = cleaned_glycosignal/total_int)

fractional_intensities %>%
  ggplot(aes(x = glycan_type, y = frac_int, colour = tissue)) +
  geom_violin() +
  ggforce::geom_sina() +
  stat_summary(aes(group = tissue),
    fun = median, geom = "crossbar",
    position = position_dodge(0.9), colour = "black"
  ) +
  tilted


fractional_intensities %>% 
  group_by(tissue, glycan_type) %>% 
  summarise(sum_int = log2(sum((cleaned_glycosignal)))) %>% 
  ggplot(aes(x = glycan_type, fill = tissue, y = sum_int)) +
  geom_bar(stat = "identity", pos = "dodge")+
  tilted +
  labs(y = "log2 summed\n intensity")
```


```{r}
fractional_intensities %>% 
  acast(mouse + Observed.Modifications ~ tissue, value.var = "total_int") %>% 
  ComplexHeatmap::Heatmap(show_row_names = F)
```


```{r}
lowerFn <- function(data, mapping, method = "lm", ...) {
  p <- ggplot(data = data, mapping = mapping) +
    geom_point(colour = "black", size = 0.5) +
    #geom_smooth(method = method, color = "red", ...)  +
    geom_abline(slope =1, colour = "red")
  p
}

fractional_intensities %>% 
  mutate(cleaned_glycosignal = log2(cleaned_glycosignal)) %>% 
  dcast(mouse + Observed.Modifications ~ tissue, value.var = "cleaned_glycosignal") %>% 
  GGally::ggpairs(columns = c(3, 4, 5),  lower = list(continuous = GGally::wrap(lowerFn, method = "lm") ))
```


# Subcellular

```{r}
locations_of_interest <- c("Endoplasmic reticulum", "Golgi apparatus","Lysosome", "Plasma membrane", "Extracellular")

compartment_anno <- tt <- read_tsv("comparison_to_datasets/combined_loc_df.tsv")%>%
  mutate(description = ifelse(grepl("Extracellular", description), "Extracellular", description)) %>% 
  filter(description %in% locations_of_interest) %>% 
  group_by(gene) %>%
  filter(score > 2.5) %>%
  mutate(species = ifelse(grepl("MUS", ensembl_prot), "mus musculus", "homo sapiens")) %>% 
  rename(Gene = gene) %>% 
  group_by(Gene) %>% 
  mutate(n_annos =n()) 

pm_proteins <- compartment_anno %>% 
  filter(description == "Plasma membrane")

extracell_proteins <- compartment_anno %>% 
  filter(description == "Extracellular")

compartment_anno <- compartment_anno %>% 
  mutate(description = ifelse(Gene %in% extracell_proteins$Gene, "Extracellular", description)) %>% 
  mutate(description = ifelse(Gene %in% pm_proteins$Gene, "Plasma membrane", description)) %>% 
  distinct(species, Gene, description)

res_mouse_tissue_atlas$corr_per_site_per_mouse %>%
  filter(tissue1 != tissue2)%>% 
  left_join(compartment_anno %>%  select(Gene, description), by = "Gene") %>% 
  drop_na() %>% 
  ggplot(aes(x = value, colour = description)) +
  geom_density(size = 0.8) +
  scale_color_brewer(palette =  "Set2")

res_mouse_tissue_atlas$corr_per_site_per_mouse %>%
  filter(tissue1 != tissue2)%>% 
  left_join(compartment_anno %>%  select(Gene, description), by = "Gene") %>% 
  drop_na() %>% 
  mutate(description = factor(description, levels = c("Endoplasmic reticulum", "Golgi apparatus","Lysosome", "Plasma membrane", "Extracellular"))) %>% 
  ggplot(aes(y = value, x = description, colour = description)) +
  geom_violin() +
  geom_boxplot() +
  scale_color_brewer(palette =  "Set2") +
  tilted
```



# Limma

### Input


```{r}
res_mouse_tissue_atlas$eset <- ttt <-  res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  distinct(Gene, Protein.ID, n_position,Observed.Modifications, glycoID,sample, tissue, mouse, replicate, cleaned_glycosignal) %>% 
  drop_na() %>% 
  acast(glycoID ~ tissue+mouse+replicate, value.var = "cleaned_glycosignal", fun.aggregate = median) 
  # acast(glycoID ~ tissue+mouse+replicate, 
  #       value.var = "cleaned_glycosignal", fun.aggregate = median) 
  # acast(Gene +Protein.ID + n_position ~ tissue+mouse+replicate, 
  #       value.var = "cleaned_glycosignal", fun.aggregate = median) 

res_mouse_tissue_atlas$eset <- tt <- res_mouse_tissue_atlas$eset[complete.cases(res_mouse_tissue_atlas$eset),]
res_mouse_tissue_atlas$eset %>%  boxplot()

```

### Design

```{r}
t_ <- str_replace(colnames(res_mouse_tissue_atlas$eset), "_mouse\\d_rep\\d", "")
g_ <- str_replace(colnames(res_mouse_tissue_atlas$eset), "_rep\\d", "")
r_ <- str_extract(colnames(res_mouse_tissue_atlas$eset), "rep\\d")
m_ <- str_extract(colnames(res_mouse_tissue_atlas$eset), "mouse\\d")

res_mouse_tissue_atlas$design <- model.matrix(~ 0 + g_+ r_ )
res_mouse_tissue_atlas$design

res_mouse_tissue_atlas$contr <-
  cbind(
    "A_brain_mouse1_\nvs_brain_mouse2" = c(1, -1, rep(0, 6)),
    "B_liver_mouse1_\nvs_liver_mouse2" = c(rep(0, 4), 1, -1, 0, 0),
    "C_kidney_mouse1_\nvs_kidney_mouse2" = c(0,0, 1, -1, rep(0, 4))
    # "D_brain_vs_other" = c(0.5,0.5, rep(-1/4, 4), 0, 0),
    # "E_liver_vs_other" = c(rep(-1/4, 4), 0.5, 0.5, 0, 0),
    # "F_kidney_vs_other" = c(rep(-1/4, 2), 0.5, 0.5,rep(-1/4, 2), 0, 0)
  )
res_mouse_tissue_atlas$contr
```

### Fit

```{r}
fit <- lmFit(res_mouse_tissue_atlas$eset, res_mouse_tissue_atlas$design)
fit <- contrasts.fit(fit, res_mouse_tissue_atlas$contr)
fit <- eBayes(fit)
```

### Extract results

```{r}

# res_mouse_tissue_atlas$limma_results <- tt <-  lapply(colnames(fit$coefficients), function(x) {
#   limma::topTable(fit, coef = x, number = Inf) %>%
#     mutate(contrast = x, dataset = "glyco") %>%
#     rownames_to_column(var = "ID")
# }) %>%
#   bind_rows() %>%
#   separate(
#     ID, 
#            into = c("Gene", "Protein.ID", "n_position"), 
#            sep = "_", 
#            remove = F
#     ) %>%
#   #separate(ID, into = c("Protein.ID", "n_position"), sep = "_", remove = F) %>%
#   mutate(
#     hit = ifelse((logFC > log2(2) & adj.P.Val < 0.05), "hit", "no hit"),
#     hit = ifelse((logFC < log2(1 / 2) & adj.P.Val < 0.05), "hit", hit)
#   ) %>%
#   mutate(
#     direction = ifelse(logFC > 0 & hit == "hit", "up", "no change"),
#     direction = ifelse(logFC < 0 & hit == "hit", "down", direction)
#   ) %>%
#   mutate(ID = ifelse(is.na(ID), Protein.ID, ID))

res_mouse_tissue_atlas$limma_results <- tt <-  lapply(colnames(fit$coefficients), function(x) {
  limma::topTable(fit, coef = x, number = Inf) %>%
    mutate(contrast = x, dataset = "glyco") %>%
    rownames_to_column(var = "ID")
}) %>%
  bind_rows() %>%
  separate(ID, into = c("Protein.ID", "Gene", "Modified.Peptide", "Observed.Modifications","glycan_type", "n_position"), sep = "_", remove = F) %>%
  #separate(ID, into = c("Protein.ID", "n_position"), sep = "_", remove = F) %>%
  mutate(
    hit = ifelse((logFC > log2(2) & adj.P.Val < 0.05), "hit", "no hit"),
    hit = ifelse((logFC < log2(1 / 2) & adj.P.Val < 0.05), "hit", hit)
  ) %>%
  mutate(
    direction = ifelse(logFC > 0 & hit == "hit", "up", "no change"),
    direction = ifelse(logFC < 0 & hit == "hit", "down", direction)
  ) %>%
  mutate(ID = ifelse(is.na(ID), Protein.ID, ID))

# res_mouse_tissue_atlas$limma_results <- tt <-  lapply(colnames(fit$coefficients), function(x) {
#   limma::topTable(fit, coef = x, number = Inf) %>%
#     mutate(contrast = x, dataset = "glyco") %>%
#     rownames_to_column(var = "ID")
# }) %>%
#   bind_rows() %>%
#   separate(ID, into = c("Observed.Modifications"), sep = "_", remove = F) %>%
#   #separate(ID, into = c("Protein.ID", "n_position"), sep = "_", remove = F) %>%
#   mutate(
#     hit = ifelse((logFC > log2(1.5) & adj.P.Val < 0.05), "hit", "no hit"),
#     hit = ifelse((logFC < log2(1 / 1.5) & adj.P.Val < 0.05), "hit", hit)
#   ) %>%
#   mutate(
#     direction = ifelse(logFC > 0 & hit == "hit", "up", "no change"),
#     direction = ifelse(logFC < 0 & hit == "hit", "down", direction)
#   ) %>%
#   mutate(ID = ifelse(is.na(ID), Protein.ID, ID))
```

```{r, fig.height=10, fig.width =10}
n_mod <-  res_mouse_tissue_atlas$glyco_psm_corrected %>%
  drop_na() %>% 
  distinct(glycoID, Observed.Modifications) %>% 
  group_by(Observed.Modifications) %>% 
  count()

ggplot(
  res_mouse_tissue_atlas$limma_results,
  aes(x = logFC, y = -log10(P.Value), colour = direction)
) +
  geom_point(alpha = 0.5, stroke = 0) +
  #geom_text_repel(aes(label = Gene), size =2) +
  scale_colour_manual(values = c("dodgerblue4", "lightgrey", "indianred")) +
  facet_wrap(~contrast) +
  # theme_bw(14) +
  theme(panel.grid = element_blank()) +
  labs(x = "log2 fold-change", y = "-log10 p-value")

ggplot(
  res_mouse_tissue_atlas$limma_results,
  aes(x = logFC, y = -log10(P.Value), colour = direction)
) +
  geom_point(alpha = 0.5, stroke = 0) +
  #geom_text_repel(aes(label = Gene), size =2) +
  scale_colour_manual(values = c("dodgerblue4", "lightgrey", "indianred")) +
  facet_grid(glycan_type~contrast) +
  # theme_bw(14) +
  theme(panel.grid = element_blank()) +
  labs(x = "log2 fold-change", y = "-log10 p-value")
```

```{r}
int <- "P97449_Anpep_n[305]NPNN[2116]NTIHPNLR_HexNAc(3)Hex(5)NeuAc(2) % 2001.6932_sialylated_784"

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(glycoID == int) %>% 
  ggplot(aes(x = tissue, y = cleaned_glycosignal)) +
  geom_point()

int <- "Q60625_Icam5_RN[3197]GTQRGLR_HexNAc(8)Hex(9) % 3083.1105_complex/hybrid_74"

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(glycoID == int) %>% 
  ggplot(aes(x = tissue, y = cleaned_glycosignal)) +
  geom_point()

int <- "P28843_Dpp4_n[305]LDFIVLN[1988]ETR_HexNAc(3)Hex(6)Fuc(2) % 1873.6709_fucosylated_514"

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(glycoID == int) %>% 
  ggplot(aes(x = tissue, y = cleaned_glycosignal)) +
  geom_point()

int <- "HexNAc(7)Hex(5)Fuc(2)NeuGc(2) % 3138.1163"

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(Observed.Modifications == int) %>% 
  ggplot(aes(x = tissue, y = cleaned_glycosignal)) +
  geom_violin() +
  geom_point() +
  geom_boxplot() 
  facet_wrap(~glycoID)

res_mouse_tissue_atlas$glyco_psm_corrected %>% 
  filter(Gene == "Cd47") %>% 
  ggplot(aes(x = Observed.Modifications, y = (cleaned_glycosignal), colour = tissue)) +
  ggforce::geom_sina() +
  facet_wrap(~n_position, scales = "free") 
  tilted
```


# Enrichment

```{r}
library(STRINGdb)
string_db <- STRINGdb$new( version="11.5", species=10090,
                           score_threshold=200, network_type="full", input_directory="")

example1_mapped <- res_mouse_tissue_atlas$glyco_psm %>%  ungroup()  %>%  distinct(Gene) %>% as.data.frame()
example1_mapped <- string_db$map(example1_mapped, "Gene", removeUnmappedRows = TRUE )

string_db <- STRINGdb$new( version="11.5", species=10090,
                           score_threshold=200, network_type="full", input_directory="",
                           backgroundV =  unique(example1_mapped$STRING_id))
```

```{r}
res_mouse_tissue_atlas$ER_limma_liver <- tt<-  res_mouse_tissue_atlas$limma_results %>%
  filter(hit == "hit" & contrast == "B_liver_mouse1_\nvs_liver_mouse2") %>% 
  distinct(contrast, Gene) %>% 
  filter(toupper(Gene) %in% example1_mapped$Gene) %>% 
  drop_na() %>% 
  group_by(contrast) %>% 
  nest() %>%
  mutate(ER = map(data, function(df) {
    ER = string_db$get_enrichment(df$Gene)
    if(nrow(ER) >0){
       ER = as.data.frame(ER)
    } else {
      ER = data.frame(category = "dummy")
    }
  })) %>%
  select(-data) %>%
  unnest(cols = c(ER))


res_mouse_tissue_atlas$ER_limma_liver %>% 
  filter(!grepl("PMID", category)) %>% 
  group_by(category) %>% 
  slice_min(n=5, order_by = fdr) %>% 
  ggplot(aes(x= -log10(fdr), y= reorder(description, fdr), size = number_of_genes, colour = category)) +
  geom_vline(xintercept = -log10(0.01)) +
  geom_point() +
  theme_bw()
```



```{r}
library(biomaRt)
ensembl <- useEnsembl(biomart = "genes", dataset = "mmusculus_gene_ensembl")

ipro = getBM(attributes=c("uniprot_gn_id", "interpro","interpro_description","interpro_short_description", "interpro_start", "interpro_end"), 
             filters = "uniprot_gn_id",
             values= unique(res_mouse_tissue_atlas$glyco_psm$Protein.ID), 
             mart=ensembl)

```

```{r, fig.height=6, fig.width=7}
hit_def <- res_mouse_tissue_atlas$limma_results %>% 
  filter(contrast == "B_liver_mouse1_\nvs_liver_mouse2" & hit == "hit") %>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  distinct(Protein.ID, siteID)

hit_def <- res_mouse_tissue_atlas$limma_results %>% 
  filter(contrast == "B_liver_mouse1_\nvs_liver_mouse2") %>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  distinct(Protein.ID, siteID) %>% 
  mutate(hit = ifelse(siteID %in% hit_def$siteID, "hit", "no hit"))

res_mouse_tissue_atlas$domain_ORA_limma <- tt <- hit_def %>% 
  mutate(mouse = "mouse") %>% 
  ungroup() %>% 
  separate(siteID, into = c("Gene", "n_position"), sep = "_") %>% 
  inner_join(ipro,by = c("Protein.ID"="uniprot_gn_id"), multiple = "all") %>% 
  dplyr::rename(corr_hit = hit) %>% 
  mutate(n_position = as.numeric(n_position)) %>% 
  mutate(in_domain = ifelse(n_position> interpro_start & n_position < interpro_end, TRUE, FALSE)) %>% 
  filter(!is.na(interpro_description) & !is.na(interpro_start)) %>% 
  select(mouse, Protein.ID, corr_hit, interpro_description, in_domain, n_position)%>% 
  filter(in_domain == T) %>% 
  mutate(siteID = paste0(Protein.ID, "_", n_position)) %>% 
  mutate(n_sites = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit) %>% 
  mutate(n_glycan = length(unique(siteID))) %>% 
  group_by(mouse, interpro_description) %>% 
  mutate(n_domain = length(unique(siteID))) %>% 
  group_by(mouse, corr_hit, interpro_description) %>% 
  mutate(n_type_domain = length(unique(siteID))) %>% 
  ungroup() %>% 
  distinct(mouse, corr_hit, interpro_description, n_sites,n_glycan, n_domain, n_type_domain) %>% 
  rowwise() %>% 
  mutate(FT = fisher.test(x = matrix(c(n_type_domain, n_glycan-n_type_domain, n_domain-n_type_domain, n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$p.value,
         odds_ratio = fisher.test(x = matrix(
           c(n_type_domain, 
             n_glycan-n_type_domain, 
             n_domain-n_type_domain,
             n_sites-n_glycan-n_domain+n_type_domain), nrow = 2, byrow = T))$estimate)%>% 
  mutate(adj.p.val = p.adjust(FT, method = "BH"),
         hit = ifelse(adj.p.val < 0.05 & n_type_domain > 10, "hit (adj.p < 0.05)", "no hit (adj.p > 0.05)"))

hits <- res_mouse_tissue_atlas$domain_ORA_limma  %>% 
  filter(hit == "hit (adj.p < 0.05)")

hits %>% 
  mutate(odds_ratio = ifelse(!is.finite(odds_ratio),20, odds_ratio))%>% 
  filter(n_type_domain > 20) %>% 
  ggplot(aes(y = interpro_description, x = log2(odds_ratio), shape = mouse,size = n_type_domain, colour = corr_hit)) +
  geom_point() +
  geom_vline(xintercept = c(-1.7, 0, 1.7), linetype = c(3,1,3)) +
  #scale_colour_manual(values = mycolors) +
  scale_size(range = c(3, 6)) +
  labs(y = "", title = "sites with hit/no hit in l1 vs l2", subtitle = "adj.p.val < 0.05, more than 20 obs") +
  theme_bw()

```



# Save

```{r, eval = F}
hit_def <- res_mouse_tissue_atlas$corr_per_site_per_mouse%>%
  filter(tissue1 != tissue2)%>% 
  mutate(siteID = paste0(Gene, "_", n_position)) %>%
  drop_na() %>% 
  group_by(mouse, Protein.ID, siteID) %>% 
  summarise(mean_corr = mean((value))) %>% 
  mutate(classification =  ifelse(mean_corr > 0.6691, "tissue specific", "not tissue specific"))

write_tsv(hit_def, file = "2024-01-15_mousetissues_corr_per_site_classification.tsv")
```


```{r, eval = F}
save(res_mouse_tissue_atlas, file = paste0("results/", Sys.Date(), "_res_mouse_tissue_atlas.RData"))
```
